# library
```{r setup}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	include = FALSE
)

# !! IMPORTANT SET UP NOTE if you're using Web Rstudio !!
# If you plan to use generate any brain plot using ggseg() and their subsequent packages please remember to load gdal/3.7.0-icl as additional module

# For code reproducibility: and git
library(here) # project-based relative path finding
library(glue) # refer to private path

# For data manipulation
library(tidyverse)
library(data.table)
library(readxl)

# For general plot generation
library(pals) # colour pallet
library(patchwork) # for combining multiple graphs
library(ggpubr)

MRI_pheno_all.colors <- setNames(trubetskoy(13), MRI_pheno_all.list_full) # Set unique colour identifier for each MRI phenotypes
MRI_pheno.order = c(DTI_pheno.list, STR_pheno.list, "Vol")

MRI_pheno_all.full_name <- c(
  SA    = "Surface area",
  CT    = "Cortical thickness",
  FI    = "Folding index",
  GC    = "Gaussian curvature",
  GMV   = "Grey matter volume",
  IC    = "Intrinsic curvature index",
  MC    = "Mean curvature",
  FA    = "Fractional anisotropy",
  ICVF  = "Intracellular volume fraction",
  ISOVF = "Isotropic volume fraction",
  MD    = "Mean diffusivity",
  OD    = "Orientation dispersion index",
  Vol   = "Volume"
)

master_path_file = fread(here("script_manuscript/master_GWAS_file_path.txt")) # A Tab seperated path collection of all GWAS-related files, DO NOT UPLOAD TO GIT
# write.table(UPDATE, here("script_manuscript/master_GWAS_file_path.txt"), sep = "\t", quote = FALSE, col.names = FALSE)

sup_table_name_cleaner = function(df, col_name = "GWAS_name") { # remove naming common problem from files for this project
  df = df %>%
    mutate(
      across(all_of(col_name), \(x) {
        x = gsub("combined_|GWAS_|DTI_|STR_|\\.p_1e_03\\.txt|_avg_scaled|X", "", x)
        x = gsub("_as_pheno\\d+", "", x)
        x = gsub("SubCorVol", "SubCor_Vol", x)
        x = gsub("CC_", "CC.", x)
        x = gsub("Mid_", "Mid.", x)
        x = gsub("_l", ".l", x)
        x = gsub("_r", ".r", x)
        x
      })
    )
  return(df)
}

sup_table_formatter = function(df, col_name = "GWAS_name"){
  # Splits GWAS_name (e.g. "regional_CT_bankssts.lh_GWAS") into MRI_scale, MRI_pheno, and region columns while removing all other non-important info.
  df = df %>% separate_wider_delim(all_of(col_name), "_", 
                                    names = c("MRI_scale", "MRI_pheno", "region", NA, NA), 
                                    too_few = "align_start", too_many = "merge")
  
  # Sets region to NA for global phenotypes as they have no regional parcellation.
  df$region = ifelse(df$MRI_scale == "global", NA, df$region)
  
  return(df)
}

# auto generator for hpc-slurm-based commands in .sh files.
source(here("script_manuscript/slurm_scheduler.R"))

# For UKB:
# ldsc h2 and sumstat only includes autosomal HapMap3 SNP list I believe so the numbers for these files for UKB before and after adding X chromosome should not change.
# I also did check the rg between UKB_all and UKB_auto and it returned rg as 1 so I'm quite confident this is the case.
```
```{r template for master path update, eval=FALSE, include=FALSE}
UKB_manhattan = list.files(here("meta_MRI_post_GWAS/UKB_manhattan"), full.names = TRUE)
UKB_p_1e03.df = data.frame(GWAS_name = basename(UKB_manhattan), UKB_all_GWAS_path.P1e03 = UKB_manhattan)
UKB_p_1e03.df$GWAS_name = gsub("combined_GWAS_|DTI_|STR_|.p_1e_03.txt|_avg_scaled|X", "", UKB_p_1e03.df$GWAS_name)
UKB_p_1e03.df$GWAS_name = gsub("SubCorVol", "SubCor_Vol", UKB_p_1e03.df$GWAS_name)
UKB_p_1e03.df$GWAS_name = gsub("CC_", "CC.", UKB_p_1e03.df$GWAS_name)
UKB_p_1e03.df$GWAS_name = gsub("_l", ".l", UKB_p_1e03.df$GWAS_name)
UKB_p_1e03.df$GWAS_name = gsub("_r", ".r", UKB_p_1e03.df$GWAS_name)
UKB_p_1e03.df$GWAS_name = gsub("Mid_", "Mid.", UKB_p_1e03.df$GWAS_name)

test = full_join(master_path_file,UKB_p_1e03.df)
write.table(test, here("script_manuscript/master_GWAS_file_path.txt"), sep = "\t", quote = FALSE, row.names = FALSE)
```
```{r Function: ldsc batch read h2 log}
ldsc_h2_batch_read <- function(file_paths, file_names=NULL) {
  # Initialize an empty list to store results
  results_list <- vector("list", length(file_paths))
  
  # Loop through each file path
  for (i in seq_along(file_paths)) {
    current_path <- file_paths[i]
    
    # Skip if NA, empty, or file doesn't exist
    if (is.na(current_path) || current_path == "" || !file.exists(current_path)) next
    
    current_file <- readLines(current_path, warn = FALSE)  # Avoid warnings for empty lines
    
    # Safe function to extract numeric values
    extract_value <- function(pattern, word_index) {
      pos <- grep(pattern, current_file)
      if (length(pos) > 0) { # check if pattern was found
        value <- scan(current_path, what = "character", skip = pos[1] - 1, nlines = 1, quiet = TRUE)
        if (length(value) >= word_index) { # check if word index exist
          return(as.numeric(gsub("[()]", "", value[word_index])))
        }
      }
      return(NA)  # Return NA if not found
    }
    
    # Extract values
    h2 <- extract_value("Total Observed scale h2", 5)
    h2_se <- extract_value("Total Observed scale h2", 6)
    
    lambda_GC <- extract_value("Lambda GC", 3)
    mean_chi_squared <- extract_value("Chi", 3)
    
    Intercept <- extract_value("Intercept", 2)
    Intercept_se <- extract_value("Intercept", 3)
    
    ratio <- extract_value("Ratio", 2)
    ratio_se <- extract_value("Ratio", 3)
    
    # If file names were provided accordingly
    if (is.null(file_names)) {GWAS_name = basename(current_path)} else {GWAS_name = file_names[i]}
    
    # Store results in the list
    results_list[[i]] <- data.frame(
      GWAS_name = GWAS_name,
      Lambda_GC = lambda_GC,
      Mean_chi_square = mean_chi_squared,
      h2 = h2,
      h2_se = h2_se,
      Intercept = Intercept,
      Intercept_se = Intercept_se,
      Ratio = ratio,
      Ratio_se = ratio_se,
      stringsAsFactors = FALSE
    )
  }
  
  # Combine all results into a single data frame
  results_df <- do.call(rbind, results_list)
  
  return(results_df)
}
```
```{r Function: ldsc batch read rg log}
ldsc_rg_batch_read = function(file_path, file_name){
  require(tidyverse)
  df_list = vector("list", length(file_path))
  
  for (i in seq_along(file_path)){
    # read file
    current_path = file_path[i]
    
    # read as vector of lines
    current_file = readLines(current_path)
    
    # Find the position of "Summary of Genetic Correlation Results" for current_file
    result_position = grep("Summary of Genetic Correlation Results", current_file)
    start_position = result_position + 1
    end_position = length(current_file) - 3
    
    # read ldsc table
    current_file_formatted = current_file[start_position:end_position]
    
    df = fread(text = current_file_formatted)
    
    # format for easy visualisation
    df$p1_base = basename(df$p1)
    df$p1_base = gsub(".sumstats|.gz", "", df$p1_base)
    
    df$p2_base = basename(df$p2)
    df$p2_base = gsub(".sumstats|.gz", "", df$p2_base)
    
    # Find p label - if "<" detected assign *** directly, otherwise cut as normal
    df$p.label <- ifelse(grepl("<", df$p), "***",
                    as.character(cut(as.numeric(df$p), 
                        breaks = c(-Inf, 0.001, 0.01, 0.05, Inf), 
                        labels = c("***", "**", "*", NA))))
    
    # Find Confidence Interval (CI) at 95%
    df$CI <- df$se*1.96
    
    df_list[[i]] = df
  }
  
  # Merge all data frames into one
  result = Reduce(function(x, y) full_join(x, y), df_list)
  result = result %>% relocate(p1_base, p2_base) %>% select(!c(p1, p2))
  return(result)
}
```
```{r Function: PLINK batch read clump files}
clump_batch_read <- function(GWAS_name, GWAS_path) {
  # check for NA as some do not have loci pass 5e-8 threshold or simply not available for the GWAS selected
  valid_paths <- !is.na(GWAS_path) & GWAS_path != "" & file.exists(GWAS_path)
  
  # Create a list of data frames with the GWAS_name column
  gwas_tables <- map2(GWAS_name[valid_paths], GWAS_path[valid_paths], ~ {
    fread(.y) %>% 
      mutate(GWAS_name = .x)
  })
  # Merge all tables using full_join
  merged_table <- reduce(gwas_tables, full_join)
  
  return(merged_table)
}
```

# Section 1: Basic GWAS info
```{r Section 1 specific library}
# For brain plot generation
library(ggseg)
library(ggseg.formats)
library(ggsegGlasser)
```

## Supplement table 1: UKB and ABCD demographic info
```{r ST1: Sex and Age Information in regards to UKB and ABCD used for GWAS generation}
UKB_demographic = fread(here("MRI_GWAS_COVAR/components/continous.COVAR.txt"))
UKB_demographic$age_years = UKB_demographic$age_days/365
UKB_demographic = UKB_demographic %>% mutate(SEX = recode(gSEX, `0` = "Female", `1` = "Male"))

table(UKB_demographic$SEX)
summary(round(UKB_demographic$age_years, digits = 2))
round(sd(UKB_demographic$age_years), digits = 2)

# Grab ABCD.COVAR.ALL from ABCD_data_preprocessing.Rmd
ABCD.PCA = fread(here("MRI_GWAS_COVAR/components/ABCD_PCsforGWAS_European.txt"))
ABCD.COVAR = fread(glue("{path_vw}ABCD/Imaging_phenotypes/ABCD_allcovariates.txt"))
ABCD.COVAR.filtered = left_join(ABCD.PCA, ABCD.COVAR, by = c("Sample_name" = "FID"))

ABCD_demographic = ABCD.COVAR.filtered %>% select(gender, interview_age)
ABCD_demographic.noNA = na.omit(ABCD_demographic) # remove anyone who do not have gender and age
ABCD_demographic.noNA$age_years = ABCD_demographic.noNA$interview_age/12
ABCD_demographic.noNA = ABCD_demographic.noNA %>% mutate(SEX = recode(gender, `1` = "Female", `0` = "Male"))

table(ABCD_demographic.noNA$SEX)
summary(round(ABCD_demographic.noNA$age_years, digits = 2))
round(sd(ABCD_demographic.noNA$age_years), digits = 2)
```
## Supplement file 1: MRI phenotype correlation matrix
```{r SF1: Generate MRI phenotype correlation matrix, eval=FALSE, include=FALSE}
# Phenotype Spectral Decomposition (PhenoSpD) used for phenotype correlation estimation and multiple testing correction for GWAS: https://github.com/MRCIEU/PhenoSpD
# option 1 is used when we have GWAS but do not have phenotype correlation matrix, option 2 is used when we do have phenotype correlation matrix.
# This tool is based on MatSpD: Newest version and instructions avaliable at https://research.qut.edu.au/sgel/software/matspd-local-version/

# Read formatted global and regional phenotypes
file.loc = list.files(here("GWAS_Trait/imaging_data_v3"), pattern=".sd.mad.QCed.txt", full.names=TRUE)
file.name = gsub(".formatted|.sd|.mad|.QCed|.txt", "", basename(file.loc)) # remove anything fit the patterns

# for loop to read and format all MRI phenotypes
for (i in 1:length(file.loc)) { 
  # read table, drop FID column but keep IID for merging tables
  tmp_file = fread(file.loc[i], drop = c("FID")) 
  
  # format for easier reading
  colnames(tmp_file) = gsub("_avg_scaled", "", colnames(tmp_file))
  
  # rename non-IID columns (which = colnames(tmp_file)[-1]) for distinguish names
  if (grepl("regional|tract", file.loc[i])) {
    col_prefix = gsub("regional_|tract_", "", file.name[i])
    colnames(tmp_file)[-1] = paste0(col_prefix, "_", colnames(tmp_file)[-1])
  }
  assign(file.name[i], tmp_file)
}

# merge table
df_list = mget(file.name)
MRI_phenotypes = df_list %>% reduce(full_join, by='IID') %>% select(-IID)

# Generate Correlation Matrix
MRI_phenotypes = as.data.table(MRI_phenotypes) # Optimise for running times 
MRI_cor = cor(MRI_phenotypes, method = "pearson", use="pairwise.complete.obs")

write.table(MRI_cor, file = here("result_manuscript/SF1_MRI_phenotype_correlation.matrix"), sep = " ") # This keeps row and column names so it's easier to reuse later.
write.table(MRI_cor, file = glue("{path_yg}result_supplement/phenoSpD/correlationORcovariance.matrix"), sep = " ", row.names = FALSE, col.names = FALSE) 
# write in personal hpc work folder in required format for script running.
# NOTE. DO NOT CHANGE THE LINE ABOVE, THE file name, sep, row.names and col.names FORMAT WAS AS REQUIRED AS IT IS FOR THE MatSpD PROGRAMME TO RUN!!!
```
## Supplement table 2: Method Used to set up multiple testing correction threshold - Matrix Decomposition through Phenotype Spectral Decomposition (PhenoSpD)
```{r ST2: Run PhenoSpD option 2, eval=FALSE, include=FALSE}
system(glue("cd {path_vw.toolbox}PhenoSpD && Rscript ./script/phenospd.r --phenocorr {path_yg}result_supplement/phenoSpD/correlationORcovariance.matrix --out {here()}/result_manuscript/ST2_PhenoSpD_log"))
```
```{r ST2: Run MatSpD for replication, eval=FALSE, include=FALSE}
source(glue("{path_yg}result_supplement/phenoSpD/matSpDliteNewAbsMatrix.R"), chdir = TRUE)

# Check result at matSpDliteNewAbsMatrix.out and copy to result folder, renamed as ST2_MatSpD_log.txt

# Therefore, for our GWAS files, Experimental-wide significance threshold should be:
# α_EWS = α_GWS / number_of_tests = 5 × 10^-8 / 1075.29 = 4.65 × 10^-11
```
Therefore, we have 1075.29 effective independent phenotypes.
Therefore, our Experimental-wide threshold for multiple testing for our GWAS should be: α_EWS = 5 × 10⁻⁸ / 1075.29 = 4.65 × 10⁻¹¹
## Supplement table 3: Information on UKB and meta-GWAS: ldsc h2
```{r ST3: master h2 table, eval=FALSE, include=FALSE}
meta_GWAS_h2 = ldsc_h2_batch_read(file_names = master_path_file$GWAS_name, file_paths = master_path_file$meta_h2_path)
UKB_GWAS_h2 = ldsc_h2_batch_read(file_names = master_path_file$GWAS_name, file_paths = master_path_file$UKB_h2_path)

# separate name into three different columns for name and visualisation
meta_GWAS_h2 = sup_table_formatter(meta_GWAS_h2)

# separate name into three different columns for name and visualisation
UKB_GWAS_h2 = sup_table_formatter(UKB_GWAS_h2)

# full_join for comparison
meta_col_name = colnames(meta_GWAS_h2)
meta_col_name[(4:length(meta_col_name))] = paste0("meta_", meta_col_name[(4:length(meta_col_name))])
colnames(meta_GWAS_h2) = meta_col_name

UKB_col_name = colnames(UKB_GWAS_h2)
UKB_col_name[(4:length(UKB_col_name))] = paste0("UKB_", UKB_col_name[(4:length(UKB_col_name))])
colnames(UKB_GWAS_h2) = UKB_col_name

master_GWAS_h2 = full_join(UKB_GWAS_h2, meta_GWAS_h2)
write.table(master_GWAS_h2, here("result_manuscript/ST3_UKB_and_meta_h2.txt"), quote = F, row.names = F)
```
## Statistical Tests: MRI and Coritcal location Heritability Differences
```{r heritability difference between MRI pheno, MRI type and MRI scale}
master_GWAS_h2 = fread(here("result_manuscript/ST3_UKB_and_meta_h2.txt"))

# Add tag for whether it's from sMRI or dMRI
master_GWAS_h2$MRI_type = ifelse(master_GWAS_h2$MRI_pheno %in% c(STR_pheno.list, "Vol"), "STR", "DTI")
h2_MRI_type.t_test = t.test(UKB_h2 ~ MRI_type, data = master_GWAS_h2)

# t-test between MRI scale: cortical (global + regional) vs subcortical (tract + subcor)
master_GWAS_h2$MRI_brain_scale = ifelse(master_GWAS_h2$MRI_scale %in% c("global", "regional"), "cor", "subcor")
h2_MRI_brain_type.t_test = t.test(UKB_h2 ~ MRI_brain_scale, data = master_GWAS_h2)
```
```{r Check heritability p}
master_GWAS_h2$UKB_p = 2 * pnorm(-abs(master_GWAS_h2$UKB_h2 / master_GWAS_h2$UKB_h2_se))
table(master_GWAS_h2$UKB_p < 0.05) # 2332 so yes it's all significant

master_GWAS_h2$meta_p = 2 * pnorm(-abs(master_GWAS_h2$meta_h2 / master_GWAS_h2$meta_h2_se))
table(master_GWAS_h2$meta_p < 0.05) # 2197 is total number without tract so yes it's all significant
```
```{r average h2 across different MRI phenotypes}
master_GWAS_h2 %>% 
  group_by(MRI_pheno) %>% 
  summarise(mean_h2 = mean(UKB_h2, na.rm = TRUE)) %>%
  arrange(desc(mean_h2))
```
```{r average h2 improvement compare to previous vw2023 GWAS}
# read in vw2023 h2
MRI_pheno_all.full_name.r = c(
  setNames(names(MRI_pheno_all.full_name), MRI_pheno_all.full_name),
  "Local gyrification index" = "LGI",
  "Intrinsic curvature" = "IC",
  "Volume" = "GMV"
)

MRI_pheno_all.full_name.r = MRI_pheno_all.full_name.r[!duplicated(names(MRI_pheno_all.full_name.r), fromLast = TRUE)]

vw2023.global.h2 = read_xlsx(here("script_manuscript/NIHMS1937250-supplement-Supplementary_Tables_1-34.xlsx"), sheet = "S4", skip = 3) %>% 
  select(1:2) %>%
  rename(GWAS_name = 1, meta_h2_2023 = 2)  %>% 
  mutate(GWAS_name = MRI_pheno_all.full_name.r[GWAS_name],
    GWAS_name = paste0("global_", GWAS_name),
    meta_h2_2023 = round(meta_h2_2023, 2))
  
vw2023.regional.h2 = read_xlsx(here("script_manuscript/NIHMS1937250-supplement-Supplementary_Tables_1-34.xlsx"), sheet = "S29", skip = 3) %>% select(-1)
vw2023.regional.h2.long = vw2023.regional.h2 %>%
  mutate(across(-`Region name`, ~ as.numeric(str_remove(., "\\s*\\[.*\\]")))) %>%
  pivot_longer(cols = -`Region name`, names_to = "Phenotype", values_to = "h2") %>%
  mutate(Phenotype = MRI_pheno_all.full_name.r[Phenotype],
    Phenotype = paste0("regional_", Phenotype, "_", `Region name`)) %>%
  select(Phenotype, h2) %>% 
  rename(GWAS_name = 1, meta_h2_2023 = 2)

vw2023.h2 = sup_table_formatter(full_join(vw2023.global.h2, vw2023.regional.h2.long))

# read in master h2 table (cortical only as vw2023 only have cortical data)
master_GWAS_h2 = fread(here("result_manuscript/ST3_UKB_and_meta_h2.txt")) %>% 
  select(MRI_scale, MRI_pheno, region, UKB_h2, meta_h2) %>%
  mutate(UKB_h2 = round(UKB_h2, 2),
         meta_h2 = round(meta_h2, 2)) %>% filter(MRI_scale == "global" | MRI_scale == "regional")

h2_comparison.df = left_join(master_GWAS_h2, vw2023.h2) %>% mutate(ukb_2026_2023 = UKB_h2 - meta_h2_2023, meta_2026_2023 = meta_h2 - meta_h2_2023)
h2_comparison.df$MRI_type = ifelse(h2_comparison.df$MRI_pheno %in% STR_pheno.list, "STR", "DTI")
t.test(ukb_2026_2023 ~ MRI_type, data = h2_comparison.df)
t.test(meta_2026_2023 ~ MRI_type, data = h2_comparison.df)
```
## Figure 1 AND Figure 2: Cortical and Subcortical H2 Heritability
```{r table prep for figure 1 generation}
master_GWAS_h2 = fread(here("result_manuscript/ST3_UKB_and_meta_h2.txt"))

# generate CI for figures
master_GWAS_h2 <- master_GWAS_h2 %>%
  mutate(across(ends_with("h2_se"), ~ . * 1.96, .names = "{str_replace(.col, '_se', '_CI')}"))

# Add tag for whether it's from sMRI or dMRI
master_GWAS_h2$MRI_type = ifelse(master_GWAS_h2$MRI_pheno %in% STR_pheno.list, "STR", "DTI")

# give factor level to MRI phenotypes so everything can draw in order 
MRI_pheno.order = c(DTI_pheno.list, STR_pheno.list, "Vol")
master_GWAS_h2 = master_GWAS_h2 %>%
  mutate(MRI_pheno = factor(MRI_pheno, levels = MRI_pheno.order))

# separate into tables for easy visualisation
h2_global_cor = subset(master_GWAS_h2, master_GWAS_h2$MRI_scale=="global")
h2_regional_cor = subset(master_GWAS_h2, master_GWAS_h2$MRI_scale=="regional")
h2_subcor_tract <- subset(master_GWAS_h2, MRI_scale %in% c("tract", "SubCor"))
```
```{r Figure 1: Cortical heritability - global and regional phenotypes construction}
# Figure 1a: Global IDP h2 - Bar Graph
h2_global_vis <- ggplot(h2_global_cor, aes(x = MRI_pheno, y = UKB_h2,
                                            ymin = UKB_h2 - UKB_h2_CI, 
                                            ymax = UKB_h2 + UKB_h2_CI)) + 
  geom_errorbar(color = "black", width = 0.2, linewidth = 0.5) +
  geom_point(aes(fill = MRI_pheno), size = 3, shape = 21, stroke = 0.5, color = "black") + 
  geom_text(aes(label = round(UKB_h2, 2), y = UKB_h2 + UKB_h2_CI), vjust = -0.5, size = 3.5) +
  ylim(0, 0.45) + 
  ylab("Global IDP h2") +
  scale_fill_manual(values = MRI_pheno_all.colors, guide = "none") +
  theme_classic() +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major = element_line(color = "grey90"),
        legend.position = "none",
        strip.text = element_text(margin = margin(b = 6, t = 4))) +
  facet_grid(~MRI_type, scales = "free", space = "free_x")

# Figure 1b: Regional IDP h2 - Violin Graph
h2_reg_vio.vis <- h2_regional_cor %>% 
  ggplot(aes(MRI_pheno, UKB_h2)) + 
  geom_violin(aes(fill = MRI_pheno)) +
  geom_boxplot(width = 0.1) +
  scale_fill_manual(values = MRI_pheno_all.colors, name = "MRI Phenotypes") +
  facet_grid(~MRI_type, scale = "free", space = "free_x") + 
  xlab("") + 
  ylab("Regional IDP h2") +
  ylim(0, 0.45) +
  theme_classic() +
  theme(panel.grid.major = element_line(color = "grey90"),
        strip.text = element_blank(),
        legend.position = "bottom",
        plot.margin = margin(t = 5.5, r = 5.5, b = 0, l = 5.5)) +
  guides(fill = guide_legend(nrow = 1))

# Figure 1c: Regional IDP h2 - Glassar Brain Graph
h2_reg_brain_vis <- h2_regional_cor %>% 
  ggplot(mapping=aes(fill = UKB_h2)) + 
  geom_brain(atlas = glasser(), hemi = "left", view = c("lateral", "medial"), size = 0.1, color = "darkgrey") +
  scale_fill_viridis_c(option = "magma", na.value="grey", limits = c(0, max(master_GWAS_h2$UKB_h2)), name = "h2") +
  facet_wrap(~MRI_pheno, nrow = 3) +
  theme_void() +
  theme(legend.position = "bottom",
        plot.margin = margin(0, 0, 0, 0))

# Combine all
fig1 = (h2_global_vis / h2_reg_vio.vis / h2_reg_brain_vis) + 
  plot_layout(guides = "collect", heights = c(1, 1, 2)) +
  plot_annotation(theme = theme(legend.position = "bottom"))

ggsave(here("result_manuscript/F3.1_cortical_regional_h2.png"), plot = fig1, width = 15, height = 14, dpi = 300, units = "in")
ggsave(here("result_manuscript/F3.1_cortical_regional_h2.pdf"), plot = fig1, width = 15, height = 14, dpi = 300, units = "in")
```
```{r Figure 2 Subcortical heritability - subcorvol and tract-based phenotypes construction}
# Figure 2a: Sub-cortical Volume and tract based IDP h2 - Violin Graph
h2_subcor_vis <- h2_subcor_tract %>% 
  ggplot(aes(MRI_pheno, UKB_h2)) + 
  geom_violin(aes(fill = MRI_pheno)) +
  geom_boxplot(width = 0.1) +
  scale_fill_manual(values = MRI_pheno_all.colors, name = "MRI Phenotypes") +
  facet_grid(~MRI_type, scale = "free", space = "free_x") + 
  xlab("") + 
  ylab("Regional IDP h2") +
  ylim(0, 0.45) +
  theme_classic() +
  theme(panel.grid.major = element_line(color = "grey90"),
        strip.text = element_blank(),
        legend.position = "bottom",
        plot.margin = margin(t = 5.5, r = 5.5, b = 0, l = 5.5)) +
  guides(fill = guide_legend(nrow = 1))

# Figure 2b: Tract-based IDP h2 - Brain Graph
# Note. Rational of the label_map set up is detailed in ST4

label_map <- c(
  # ---- direct ----
  "ar.l"  = "lh.ar.bbr.prep",
  "atr.l" = "lh.atr.bbr.prep",
  "cst.l" = "lh.cst.bbr.prep",
  "ilf.l" = "lh.ilf.bbr.prep",
  "unc.l" = "lh.uf.bbr.prep",
  "ar.r"  = "rh.ar.bbr.prep",
  "atr.r" = "rh.atr.bbr.prep",
  "cst.r" = "rh.cst.bbr.prep",
  "ilf.r" = "rh.ilf.bbr.prep",
  "unc.r" = "rh.uf.bbr.prep",
  "ml.l"  = "lh.mlf.bbr.prep",
  "ml.r"  = "rh.mlf.bbr.prep",
  "mcp"   = "mcp.bbr.prep",

  # ---- proxy ----
  "cgc.l" = "lh.cbd.bbr.prep",
  "cgc.r" = "rh.cbd.bbr.prep",
  "cgh.l" = "lh.cbv.bbr.prep",
  "cgh.r" = "rh.cbv.bbr.prep",
  "ifo.l" = "lh.emc.bbr.prep",
  "ifo.r" = "rh.emc.bbr.prep",
  "ptr.l" = "lh.or.bbr.prep",
  "ptr.r" = "rh.or.bbr.prep",
  "fmi"   = "cc.genu.bbr.prep",
  "fma"   = "cc.splenium.bbr.prep",

  # ---- subdivision (encoded as multiple labels) ----
  "slf.l" = "lh.slf1.bbr.prep;lh.slf2.bbr.prep;lh.slf3.bbr.prep",
  "slf.r" = "rh.slf1.bbr.prep;rh.slf2.bbr.prep;rh.slf3.bbr.prep"
)

h2_tract <- h2_subcor_tract %>%
  filter(MRI_scale == "tract") %>%
  mutate(label = label_map[region]) %>%
  filter(!is.na(label)) %>%
  tidyr::separate_rows(label, sep = ";") %>%
  select(!region)

# plot
h2_tract_brain_vis <- h2_tract %>%
  ggplot(aes(fill = UKB_h2)) +
  geom_brain(atlas = tracula(),
             view = c("axial_2", "coronal_4"),
             size = 0.1, 
             color = "darkgrey") +
  scale_fill_viridis_c(option = "magma", na.value = "grey",
                       limits = c(0, max(master_GWAS_h2$UKB_h2)), 
                       name = "h2") +
  facet_wrap(~MRI_pheno, nrow = 2) +
  theme_void() +
  theme(legend.position = "bottom")

# Figure 2c: Subcortical-Volume-based IDP h2 - Brain Graph
label_map_subcor <- c(
  "Cerebellum.Cortex"     = "Cerebellum",
  "Thalamus.Proper"       = "Thalamus",
  "Thalamus.Proper"       = "Thalamus Proper",
  "Caudate"               = "Caudate",
  "Putamen"               = "Putamen",
  "Pallidum"              = "Pallidum",
  "Hippocampus"           = "Hippocampus",
  "Amygdala"              = "Amygdala",
  "Accumbens.area"        = "accumbens area",
  "VentralDC"             = "ventraldc",
  "vessel"                = "vessel",
  "choroid.plexus"        = "choroid plexus",
  "Brain.Stem"            = "Brain Stem",
  "Optic.Chiasm"          = "Optic Chiasm",
  "CC.Posterior"          = "cc posterior",
  "CC.Mid.Posterior"      = "cc mid posterior",
  "CC.Central"            = "cc central",
  "CC.Mid.Anterior"       = "cc mid anterior",
  "CC.Anterior"           = "cc anterior"
)

h2_subcor <- h2_subcor_tract %>%
  filter(MRI_scale == "SubCor") %>%
  mutate(region = label_map_subcor[region]) %>%
  filter(!is.na(region)) %>%
  bind_rows(
    filter(., region == "Thalamus") %>%
      mutate(region = "Thalamus Proper") # ggseg() bug, if both rows not present will only fill one side
  )

# plot
h2_subcor_brain_vis <- h2_subcor %>%
  ggplot(aes(fill = UKB_h2)) +
  geom_brain(atlas = aseg(), view = c("axial_4", "sagittal", "axial_5"),
             size = 0.1, 
             color = "darkgrey") +
  scale_fill_viridis_c(option = "magma", na.value = "grey",
                       limits = c(0, max(master_GWAS_h2$UKB_h2)), 
                       name = "h2") +
  theme_void() +
  theme(legend.position = "bottom") +
  labs(title = "Subcortical Volume") +
  theme(plot.title = element_text(hjust = 0.5, size = 10))

# Combine all
fig2 = (h2_subcor_vis / h2_subcor_brain_vis / h2_tract_brain_vis) + 
  plot_layout(guides = "collect", heights = c(2, 1, 1.75)) +
  plot_annotation(theme = theme(legend.position = "bottom"))

ggsave(here("result_manuscript/F3.2_subcortical_regional_h2.png"), plot = fig2, width = 15, height = 17.5, dpi = 300, units = "in")
ggsave(here("result_manuscript/F3.2_subcortical_regional_h2.pdf"), plot = fig2, width = 15, height = 17.5, dpi = 300, units = "in")
```
# Section 2: Comparison with previous imaging GWAS
## Supplement file 2 AND Supplement file 3: Clumped results for UKB AND Clumped results for meta
```{r SF2 AND SF3: Raw clumped results for UKB GWAS AND Raw clumped results for meta-GWAS, eval=FALSE, include=FALSE}
UKB_clump = clump_batch_read(master_path_file$GWAS_name, master_path_file$UKB_clump_path)
meta_clump = clump_batch_read(master_path_file$GWAS_name, master_path_file$meta_clump_path)

UKB_clump = sup_table_formatter(UKB_clump) %>% relocate(MRI_scale, MRI_pheno, region)
meta_clump = sup_table_formatter(meta_clump) %>% relocate(MRI_scale, MRI_pheno, region)

write.table(UKB_clump, here("result_manuscript/SF2_UKB_sig.5e-8.r2.0.1.clump.txt"), row.names = F, quote = F)
write.table(meta_clump, here("result_manuscript/SF3_meta_sig.5e-8.r2.0.1.clump.txt"), row.names = F, quote = F)
```
## Supplement file 4: sig.SNP.ino in either ABCD, UKB OR meta based cohorts
```{r construction of Supplement file 4}
# These were code used to generate ST1 tables
source("/rds/user/yg330/rds-genetics_hpc-Nl99R8pHODQ/UKB/Imaging_genetics/yg330/vw_requested_results/batch_GWAS_stat_comparison.R")
source("/rds/user/yg330/rds-genetics_hpc-Nl99R8pHODQ/UKB/Imaging_genetics/yg330/vw_requested_results/batch_GWAS_stat_comparison_plus_subcor_tract.R")
source("/rds/user/yg330/rds-genetics_hpc-Nl99R8pHODQ/UKB/Imaging_genetics/yg330/vw_requested_results/missing_meta_SNP_investigate_batch_GWAS_stat_comparison.R")

SF4 = fread(here("vw_requested_results/ST1_cortical_subcortical_UKB_ABCD_meta_stat_comparison.txt"))
SF4.meta_only_SNPs = list.files(here("vw_requested_results/ST1_temp_missing"), full.names = TRUE)
SF4.meta_only_SNPs.table = SF4.meta_only_SNPs %>% lapply(fread) %>% reduce(full_join)

write.table(SF4, here("result_manuscript/SF4_sig.SNP.across.ABCD.UKB.meta_table.txt"), quote = F, row.names = F)
```
## Statistical Tests: Concordance between UKB and ABCD
```{r concordance rate t-test}
# concordance rate: BETA ONLY
EWS.sig.UKB_in_ABCD = fread(here("result_manuscript/ST5_EWS.sig.SNP.from.UKB_compare_to_ABCD_table.txt"))
EWS.sig.UKB_in_ABCD = sup_table_formatter(EWS.sig.UKB_in_ABCD)
beta_concordance_rate = table(EWS.sig.UKB_in_ABCD$concordance)
binom.test(beta_concordance_rate[1], beta_concordance_rate[2]+beta_concordance_rate[1], p = 0.5, alternative = "greater")

# concordance rate: BETA and P
EWS.sig.UKB_in_sig.ABCD <-EWS.sig.UKB_in_ABCD %>% filter(EWS.sig.UKB_in_ABCD$concordance == "Concordant" & EWS.sig.UKB_in_ABCD$ABCD_P < 0.05)
beta_and_P_concordance_rate = table(EWS.sig.UKB_in_sig.ABCD$concordance)
binom.test(beta_and_P_concordance_rate[1], beta_concordance_rate[2]+beta_concordance_rate[1], p = 0.05, alternative = "greater")

# Concordance rate by phenotype
beta_concordance_rate.by_MRI_pheno = table(EWS.sig.UKB_in_ABCD$concordance, EWS.sig.UKB_in_ABCD$MRI_pheno)
beta_concordance_rate.by_MRI_pheno.pct <- prop.table(beta_concordance_rate.by_MRI_pheno, margin = 2)["Concordant", ] * 100

beta_concordance_rate.by_MRI_pheno.pct.df = data.frame(
  MRI_pheno = names(beta_concordance_rate.by_MRI_pheno.pct),
  concordance.pct = as.numeric(beta_concordance_rate.by_MRI_pheno.pct))

# statistical t-test to see if pct is different amongst MRI type
beta_concordance_rate.by_MRI_pheno.pct.df$MRI_type = ifelse(beta_concordance_rate.by_MRI_pheno.pct.df$MRI_pheno %in% c(STR_pheno.list, "Vol"), "STR", "DTI")
beta_concordance_rate.by_MRI_pheno.pct.df.t_test = t.test(concordance.pct ~ MRI_type, data = beta_concordance_rate.by_MRI_pheno.pct.df)
beta_concordance_rate.by_MRI_pheno.pct.df.t_test
```

## Supplement table 5: Concordance Table
```{r construction of Supplement table 5}
concor.table = fread(here("result_manuscript/SF4_sig.SNP.across.ABCD.UKB.meta_table.txt"))

EWS.sig.UKB_in_ABCD = concor.table %>% filter(!is.na(ABCD_BETA) & !is.na(UKB_BETA) & UKB_P < 4.653e-11)
write.table(EWS.sig.UKB_in_ABCD, here("result_manuscript/ST5_EWS.sig.SNP.from.UKB_compare_to_ABCD_table.txt"), quote = F, row.names = F)
```
## Supplement file 5 AND Supplement table 6: Unique loci across neuroimaging GWAS
```{r generate clump with r2 0.8, eval=FALSE, include=FALSE}
bfile_path = glue("{path_UKB.vw}Imaging_genetics/yg330/UKB_LD/fileforclumping")
awk_cmd = 'awk -F"\\t" \'{print $1, $2}\''

PLINK_clump_r20.8_command = glue('
module load plink/1.9
tail -n +2 "{here("script_manuscript/master_GWAS_file_path.txt")}" | {awk_cmd} | while read -r gwas_name gwas_path; do
  if [ -f "{here("meta_MRI_post_GWAS/UKB_clump_p5e-8_r20.8")}/${{gwas_name}}_clump_5e8_r2_0.8.clumped" ]; then
    echo "SKIPPING $gwas_name: output already exists"
    continue
  fi
  plink \\
    --bfile {bfile_path} \\
    --clump "$gwas_path" \\
    --clump-field P \\
    --clump-p1 0.00000005 \\
    --clump-p2 1 \\
    --clump-r2 0.8 \\
    --clump-kb 1000 \\
    --out "{here("meta_MRI_post_GWAS/UKB_clump_p5e-8_r20.8")}/${{gwas_name}}_clump_5e8_r2_0.8"
done
')

setwd(here("script_manuscript/UKB_clump_r20.8_log")) # set script path

submit_slurm(
  account  = account,
  job_name = "gwas_clump_r2_0.8",
  command  = PLINK_clump_r20.8_command,
  time     = "8:00:00",
  mem      = "16G",
  home_dir = here("meta_MRI_post_GWAS"),
  log_dir  = here("script_manuscript/UKB_clump_r20.8_log"),
  partition = "sapphire"
)
```
```{r ST5: generate UKB clump r20.8 table as supplement file 5}
UKB_r20.8_clump = clump_batch_read(master_path_file$GWAS_name, master_path_file$UKB_clump_r20.8_path)
write.table(UKB_r20.8_clump, here("result_manuscript/SF5_UKB_sig.5e-8.r2.0.8.clump.txt"), row.names = F, quote = F)
```
```{r read and process loci information from different neuroimaging server}
# Read in clump_r20.1 clump across GWAS and find unique loci
UKB_SF2.clump.r20.1 = fread(here("result_manuscript/SF2_UKB_sig.5e-8.r2.0.1.clump.txt"))
meta_SF3.clump.r20.1 = fread(here("result_manuscript/SF3_meta_sig.5e-8.r2.0.1.clump.txt"))
SF2_and_SF3.clump.r20.1.unique = unique(c(UKB_SF2.clump.r20.1$SNP, meta_SF3.clump.r20.1$SNP))

# Read in r20.8 clumps across GWAS
UKB_r20.8_clump = fread(here("result_manuscript/SF5_UKB_sig.5e-8.r2.0.8.clump.txt"))
UKB_r20.8.clump.unique = unique(UKB_r20.8_clump$SNP)

# For previous UKB-based paper
## For vw2023 paper
MRI_pheno_all.full_name.r <- setNames(names(MRI_pheno_all.full_name), MRI_pheno_all.full_name) # for converting full name to shortened names

MRI_pheno_all.full_name.r <- c(
  MRI_pheno_all.full_name.r,
  "Local gyrification index" = "LGI",
  "Intrinsic curvature" = "IC"
)

vw2023.global = read_xlsx(here("script_manuscript/NIHMS1937250-supplement-Supplementary_Tables_1-34.xlsx"), sheet = "S1", skip = 2)
vw2023.regional = read_xlsx(here("script_manuscript/NIHMS1937250-supplement-Supplementary_Tables_1-34.xlsx"), sheet = "S26", skip = 2)
vw2023.regional.conversion_table = fread(here("script_manuscript/NIHMS1937250_region_number_name.txt"))

vw2023.global.readable = vw2023.global %>% 
  mutate(
    # 1. Replace full phenotype name with short name
    Phenotype = MRI_pheno_all.full_name.r[Phenotype],
    
    # 3. Create GWAS_name as pheno_region
    GWAS_name = paste0("global_", Phenotype))

vw2023.regional.readable = vw2023.regional %>%
  mutate(
    # 1. Replace full phenotype name with short name
    phenotype = MRI_pheno_all.full_name.r[phenotype],
    
    # 2. Replace region number with region name
    region = vw2023.regional.conversion_table$`Region name`[
      match(region, vw2023.regional.conversion_table$`Region number`)
    ],
    
    # 3. Create GWAS_name as pheno_region
    GWAS_name = paste0("regional_", phenotype, "_", region))

vw2023 = full_join(vw2023.global.readable, vw2023.regional.readable)
vw2023.unique = unique(vw2023$SNP)

# For GWAS Catalogue: Re:Format
GWAS_catalogue_neuroimaging_measurement = read_tsv(here("script_manuscript/GWAS_catalogue_EFO_0004346_associations_export.tsv"))
GWAS_catalogue_neuroimaging_measurement = GWAS_catalogue_neuroimaging_measurement %>% filter(!grepl("Alzheimer", traitName, ignore.case = TRUE)) # remove conditioned
GWAS_catalogue_neuroimaging_measurement= GWAS_catalogue_neuroimaging_measurement %>% 
  mutate(riskAllele = ifelse(grepl("^rs", riskAllele), sub("-.*", "", riskAllele), sub("^chr", "", sub("-", "_", riskAllele)))) 
GWAS_catalogue.unique = unique(GWAS_catalogue_neuroimaging_measurement$riskAllele)
# Note. GWAS Catalogue does not hold REF allele info only ALT allele info. Anything without rsid will not be put in account, dropped 143 rows.

# For ENIGMA - CLSA-PheWeb: Tophits: 1000 SNPs with the most significant P aaosications, Consider for at least P<10e-6, only top hits within 500kb shown
ENIGMA_imaging_GWAS_tophit = fread(here("script_manuscript/ENIGMA_CLAS_PheWeb_phenotype_tab_data.csv"))
ENIGMA_imaging_GWAS_tophit.brain = ENIGMA_imaging_GWAS_tophit %>% filter(!grepl("Heart|Retinal", category, ignore.case = TRUE)) # remove non-brain-measures
ENIGMA_imaging_GWAS_tophit.brain = ENIGMA_imaging_GWAS_tophit.brain %>%
  mutate(rsid_findunique = case_when(
    grepl("\\(rs\\S+\\)", variantName) ~ sub(".*\\((rs\\S+)\\).*", "\\1", variantName),
    TRUE ~ gsub(":\\s*", ":", variantName) %>%
             gsub("\\s*/\\s*", "_", .) %>%
             gsub("\\s+", "_", .)
  )) %>% select(phenocode, rsid_findunique)

# For ENIGMA - CLSA-PheWeb: Top rsid per phenotype
ENIGMA_imaging_GWAS_phenotype = fread(here("script_manuscript/ENIGMA_CLAS_PheWeb_top_hit_phenotype_data.csv"))
ENIGMA_imaging_GWAS_phenotype.brain = ENIGMA_imaging_GWAS_phenotype %>% filter(!grepl("Heart|Retinal", category, ignore.case = TRUE)) # remove non-brain-measures
ENIGMA_imaging_GWAS_phenotype.brain = ENIGMA_imaging_GWAS_phenotype.brain %>%
  filter(pval < 5e-8) %>% # must be GWS
  mutate(rsid_findunique = case_when(
    grepl("\\(rs\\S+\\)", variantName) ~ sub(".*\\((rs\\S+)\\).*", "\\1", variantName),
    TRUE ~ gsub(":\\s*", ":", variantName) %>%
             gsub("\\s*/\\s*", "_", .) %>%
             gsub("\\s+", "_", .)
  )) %>% select(phenocode, rsid_findunique)

ENIGMA = full_join(ENIGMA_imaging_GWAS_phenotype.brain, ENIGMA_imaging_GWAS_tophit.brain)
ENIGMA.unique = unique(ENIGMA$rsid_findunique)

# BIG40
BIG40 = fread(here("script_manuscript/Big40_top_hits.tsv"))
BIG40 = BIG40 %>%
  filter(pval < 5e-8) %>%
  # grab first rsid in the string if it exist, otherwise generate variant id
  mutate(rsid_findunique = ifelse(rsids == "" | is.na(rsids), 
                                  paste0(chrom, ":", pos, "_", ref, "_", alt),
                                  sub(",.*", "", rsids) %>% trimws())) %>% select(phenocode, rsid_findunique)
BIG40.unique = unique(BIG40$rsid_findunique)

ALL.unique = unique(c(GWAS_catalogue.unique, ENIGMA.unique, vw2023.unique, BIG40.unique))
sum(!SF2_and_SF3.clump.r20.1.unique %in% ALL.unique) # 22,564
sum(!UKB_r20.8.clump.unique %in% ALL.unique) # 37,530
```
```{r generate Supplement table 6: Unique loci across top hit external neuroimaging GWAS}
# Pre-compute trait mappings
ALL.clump.unique = unique(c(SF2_and_SF3.clump.r20.1.unique, UKB_r20.8.clump.unique))

## vw2023
vw2023.trait = ALL.clump.unique %>%
  as_tibble() %>%
  rename(SNP = value) %>%
  left_join(
    vw2023 %>%
      group_by(SNP) %>%
      summarise(vw2023_traits = paste(unique(GWAS_name), collapse = ", ")),
    by = "SNP"
  )

## GWAS Catalogue
## Note: GWAS Catalogue does not hold REF allele info, only ALT allele info.
## Anything without rsid will not be accounted for (~143 rows dropped).
GWAS_cat.trait = ALL.clump.unique %>%
  as_tibble() %>%
  rename(SNP = value) %>%
  left_join(
    GWAS_catalogue_neuroimaging_measurement %>%
      filter(startsWith(riskAllele, "rs")) %>%
      group_by(riskAllele) %>%
      summarise(GWAS_cat_traits = paste(unique(traitName), collapse = ", ")),
    by = c("SNP" = "riskAllele")
  )

## ENIGMA
ENIGMA.trait = ALL.clump.unique %>%
  as_tibble() %>%
  rename(SNP = value) %>%
  left_join(
    ENIGMA %>%
      group_by(rsid_findunique) %>%
      summarise(ENIGMA_traits = paste(unique(phenocode), collapse = ", ")),
    by = c("SNP" = "rsid_findunique")
  )

## BIG40
BIG40.trait = ALL.clump.unique %>%
  as_tibble() %>%
  rename(SNP = value) %>%
  left_join(
    BIG40 %>%
      group_by(rsid_findunique) %>%
      summarise(BIG40_traits = paste(unique(phenocode), collapse = ", ")),
    by = c("SNP" = "rsid_findunique")
  )

## UKB r2<0.1
UKB_r20.1.trait = ALL.clump.unique %>%
  as_tibble() %>%
  rename(SNP = value) %>%
  left_join(
    UKB_SF2.clump.r20.1 %>%
      mutate(GWAS_name = ifelse(is.na(region), 
                                paste(MRI_scale, MRI_pheno, sep = "_"),
                                paste(MRI_scale, MRI_pheno, region, sep = "_"))) %>%
      group_by(SNP) %>%
      summarise(UKB_r20.1_traits = paste(unique(GWAS_name), collapse = ", ")),
    by = "SNP"
  )

## meta r2<0.1
meta_r20.1.trait = ALL.clump.unique %>%
  as_tibble() %>%
  rename(SNP = value) %>%
  left_join(
    meta_SF3.clump.r20.1 %>%
      mutate(GWAS_name = ifelse(is.na(region), 
                                paste(MRI_scale, MRI_pheno, sep = "_"),
                                paste(MRI_scale, MRI_pheno, region, sep = "_"))) %>%
      group_by(SNP) %>%
      summarise(meta_r20.1_traits = paste(unique(GWAS_name), collapse = ", ")),
    by = "SNP"
  )

## UKB r2<0.8
UKB_r20.8.trait = ALL.clump.unique %>%
  as_tibble() %>%
  rename(SNP = value) %>%
  left_join(
    UKB_r20.8_clump %>%
      group_by(SNP) %>%
      summarise(UKB_r20.8_traits = paste(unique(GWAS_name), collapse = ", ")),
    by = "SNP"
  )

# Join all trait mappings onto main df
SF6.df = data.frame(
  rsid = ALL.clump.unique,
  # clump
  In_UKB_clump_r2_0.1 = ALL.clump.unique %in% UKB_SF2.clump.r20.1$SNP,
  In_UKB_clump_r20.8 = ALL.clump.unique %in% UKB_r20.8.clump.unique,
  In_meta_clump_r2_0.1 = ALL.clump.unique %in% meta_SF3.clump.r20.1$SNP,
  # pheno
  In_vw2023 = ALL.clump.unique %in% vw2023.unique,
  In_GWAS_cat_EFO_004346 = ALL.clump.unique %in% GWAS_catalogue.unique,
  In_ENIGMA_CLSA_PheWeb_brain = ALL.clump.unique %in% ENIGMA.unique,
  IN_BIG40 = ALL.clump.unique %in% BIG40.unique) %>%
  left_join(UKB_r20.1.trait,   by = c("rsid" = "SNP")) %>%
  left_join(UKB_r20.8.trait,   by = c("rsid" = "SNP")) %>%
  left_join(meta_r20.1.trait,  by = c("rsid" = "SNP")) %>%
  left_join(vw2023.trait,      by = c("rsid" = "SNP")) %>%
  left_join(GWAS_cat.trait,    by = c("rsid" = "SNP")) %>%
  left_join(ENIGMA.trait,      by = c("rsid" = "SNP")) %>%
  left_join(BIG40.trait,       by = c("rsid" = "SNP"))
  

write.table(SF6.df, here("result_manuscript/ST6_clumped_loci_overlap_across_neuroimaging_GWAS_external.txt"), quote = F, row.names = F, sep = "\t")
```
```{r concordance between 2026 and 2023 independent loci }
SNP2023_vs_2026_concordance = SF6.df %>%
  filter(In_vw2023 == TRUE) %>%
  mutate(
    split_UKB_r20.1  = strsplit(UKB_r20.1_traits, ", "),
    split_meta_r20.1 = strsplit(meta_r20.1_traits, ", "),
    split_UKB_r20.8  = strsplit(UKB_r20.8_traits, ", "),
    split_vw2023     = strsplit(vw2023_traits, ", ")
  ) %>%
  mutate(
    any_overlap = mapply(function(a, b, c, v) 
      any(c(a, b, c) %in% v),
      split_UKB_r20.1, split_meta_r20.1, split_UKB_r20.8, split_vw2023)
  )

SNP2023_vs_2026_concordance %>% summarise(
  yes = sum(any_overlap,  na.rm = TRUE),
  no  = sum(!any_overlap, na.rm = TRUE)
)
```
## Supplement table 7: rg against previous and external GWAS
```{r rg against previous and external GWAS: vw2023, subcortical volume, tract}
rg_comparison.name = list.files(here("script_manuscript/rg_UKB_vs_past_MRI"), pattern = ".log")
rg_comparison.name = gsub(".log", "", rg_comparison.name)

rg_comparison.path = list.files(here("script_manuscript/rg_UKB_vs_past_MRI"), pattern = ".log", full.names = TRUE)
rg_comparison.table = ldsc_rg_batch_read(rg_comparison.path, rg_comparison.name) %>%filter(!grepl("_tract_MD_", p1_base))

rg_MRI_2026_vs_MRI_2023 = fread(glue("{path_yg}SSHscript/old_vs_new_ldsc/rg_global_ABCD_vs_UKB.txt"), drop = "V1")
rg_MRI_2026_vs_MRI_2023$p2_base = gsub("_as_pheno[0-9]+", "", rg_MRI_2026_vs_MRI_2023$p2_base)
rg_MRI_2026_vs_MRI_2023$p1_base = gsub("ABCD_", "", rg_MRI_2026_vs_MRI_2023$p1_base)

rg_MRI_2026_vs_MRI_2023 = rg_MRI_2026_vs_MRI_2023 %>% filter(p1_base == p2_base)
ST7 = full_join(rg_comparison.table, rg_MRI_2026_vs_MRI_2023)

write.table(ST7, here("result_manuscript/SF7_rg_comparison_across_external_and_past_MRI_GWAS.txt"), quote = F, row.names = F)
```

## Supplementary Figure 1: Concordance Rate by MRI Phenotype and genetic correlation with different external neuroimaging GWAS
```{r Supplementary Figure 1: Concordance Rate by MRI Phenotype}
concor.pval <- beta_concordance_rate.by_MRI_pheno.pct.df.t_test$p.value
concor.means <- beta_concordance_rate.by_MRI_pheno.pct.df.t_test$estimate
concor.annotate = glue(
  "t-test p = {round(concor.pval, digits = 6)}
  Mean DTI-based IDP = {round(concor.means['mean in group DTI'], 1)}% 
  Mean STR-based IDP = {round(concor.means['mean in group STR'], 1)}%")

s.fig1 = ggplot(beta_concordance_rate.by_MRI_pheno.pct.df,
       aes(x = MRI_pheno, y = concordance.pct, fill = MRI_pheno)) +
  geom_col(color = "black", linewidth = 0.3) +
  scale_fill_manual(values = MRI_pheno_all.colors, name = "MRI Phenotypes") +
  facet_wrap(~MRI_type, scales = "free_x") +
  coord_cartesian(ylim = c(50, 100)) +
  xlab("") +  
  ggtitle(concor.annotate) +
  theme_classic() +
  theme(
    legend.position = "bottom",
    panel.grid.major.x = element_blank()) +
  ylab("Concordance (%)")

p.Q = ggplot(EWS.sig.UKB_in_ABCD, aes(x = Q)) +
  geom_histogram(fill = "steelblue", color = "black", linewidth = 0.3, bins = 30) +
  theme_classic() +
  labs(title = "Distribution of Q", x = "Q", y = "Count")

p.I = ggplot(EWS.sig.UKB_in_ABCD, aes(x = I)) +
  geom_histogram(fill = "tomato", color = "black", linewidth = 0.3, bins = 30) +
  theme_classic() +
  labs(title = "Distribution of I", x = "I", y = "Count")

combined.fig = (p.Q | p.I) / s.fig1 +
  plot_layout(heights = c(1, 2)) +
  plot_annotation(tag_levels = "A")

ggsave(here("result_manuscript/SF3.1_concor_by_phenotype.png"), plot = combined.fig, width = 15, height = 8, dpi = 300, units = "in")
ggsave(here("result_manuscript/SF3.1_concor_by_phenotype.pdf"), plot = combined.fig, width = 15, height = 8, dpi = 300, units = "in")
```

# Section 3: Widespread Pleitropy and regions of specificity
Note. Some data table called in this section were generated from UKB_hyprcoloc_edited_v2.Rmd
```{r Section 3 specific library}
library(ggrepel)
```
```{r Figure 4.3.3A - Manhattan Plot on Genetic Pleitropy Across MRI Modalities}
# Read in GWAS filtered with p < 1e-5
tract.list.df = readRDS(here("script_manuscript/manhattan_component/tract.list.df.rds"))
subcor.list.df = readRDS(here("script_manuscript/manhattan_component/subcor.list.df.rds"))
global.list.df = readRDS(here("script_manuscript/manhattan_component/global.list.df.rds"))
regional.list.df = readRDS(here("script_manuscript/manhattan_component/regional.list.df.rds"))

# Condense too sig.SNPs 

MRI_GWAS_sig.snp = list(
  tract    = tract.list.df,
  subcor   = subcor.list.df,
  global   = global.list.df,
  regional = regional.list.df
) %>%
  imap_dfr(
    ~ .x %>%
      map(~ mutate(.x, CHR = as.character(CHR))) %>%
      bind_rows(.id = "GWAS_name")
  )

MRI_GWAS_sig.snp.EWS = MRI_GWAS_sig.snp %>%
  mutate(
    CHR = replace_na(CHR, "X"),
    chr_order = if_else(CHR == "X", 23L, as.integer(CHR))
  ) %>%
  filter(P < 4.65e-11) %>%
  group_by(SNP) %>%
  mutate(
    n_pheno = n_distinct(GWAS_name),
    log10P = -log10(P)
  ) %>%
  ungroup() %>%
  arrange(chr_order, POS)

# Manhattan plot stats
chr_lengths = MRI_GWAS_sig.snp.EWS %>%
  group_by(CHR, chr_order) %>%
  summarise(
    chr_len = max(POS),
    .groups = "drop"
  ) %>%
  arrange(chr_order) %>%
  mutate(
    cum_start = cumsum(lag(chr_len + 3e7, default = 0)) # 3e7 is gap number between chromosomes, if needed add up
  )

F4.3.3.df = MRI_GWAS_sig.snp.EWS %>%
  left_join(chr_lengths %>% select(CHR, cum_start), by = "CHR") %>%
  mutate(
    cum_pos = POS + cum_start,
    log10P = ifelse(log10P>200, 200, log10P)) %>%
  arrange(chr_order, POS)

chr_mids = F4.3.3.df %>%
  group_by(CHR, chr_order) %>%
  summarise(
    mid = (min(cum_pos) + max(cum_pos)) / 2,
    .groups = "drop"
  ) %>%
  arrange(chr_order)

# Plot F4.3.3
F4.3.3A = ggplot(F4.3.3.df, aes(x = cum_pos, y = log10P, colour = n_pheno)) +
  geom_point(size = 0.8, alpha = 0.8) +
  annotate("text", x = chr_mids$mid, y = 0, label = chr_mids$CHR, size = 2.5, fontface = "bold", family = "Arial") +
  scale_colour_viridis_c(option = "plasma", name = "N phenotypes") +
  scale_x_continuous(breaks = NULL) +
  scale_y_continuous(limits = c(-5, 200), name = expression(-logP)) +
  theme_classic() +
  theme(
    panel.grid = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "bottom"
  )

ggsave(
  here("result_manuscript/F3.3A_Manhattan_across_all_MRI_GWAS.pdf"),
  F4.3.3A,
  width = 15, height = 3,
  device = cairo_pdf, bg = "white")

ggsave(
  here("result_manuscript/F3.3A_Manhattan_across_all_MRI_GWAS.png"),
  F4.3.3A,
  width = 15, height = 3,
  dpi = 600, bg = "white")
```
```{r Generation of Specificity and Pleitropic Rankings}
entropy <- function(x, n_max) {
  p <- table(x) / length(x)
  round((-sum(p * log(p))) / log(n_max), 3)
}

# read and format table
LD_summary = fread("/rds/user/yg330/hpc-work/pgs_img_2/Rscript/hyprcoloc/GWAS_coloc_LD_block_list_v2.txt") %>%
  rowwise() %>%
  mutate(# reformatt GWAS.list to approperiate supplementary table format
    traits = list(
      str_split(GWAS_list, ",")[[1]] %>%
        gsub("SubCorVol", "SubCor_Vol", ., fixed = TRUE) %>%
        gsub("CC_", "CC.", ., fixed = TRUE) %>%
        gsub("_l", ".l", ., fixed = TRUE) %>%
        gsub("_r", ".r", ., fixed = TRUE) %>%
        gsub("Mid_", "Mid.", ., fixed = TRUE)),
    
    scale = list(str_extract(unlist(traits), "^[^_]+")),
    pheno = list(str_extract(str_remove(unlist(traits), "^[^_]+_"), "^[^_]+")),
    region = list(str_remove(unlist(traits), "^[^_]+_[^_]+_")),
    
    n_traits = length(unlist(traits)),
    n_scale = n_distinct(unlist(scale)),
    n_pheno_type = n_distinct(unlist(pheno)),
    n_region = n_distinct(unlist(region)),
    
    scale_entropy = entropy(unlist(scale), n_max = 4),
    pheno_entropy = entropy(unlist(pheno), n_max = 13),
    region_entropy = entropy(unlist(region), n_max = 233)) %>%
  ungroup() %>%
  select(chr, start, end, locus, GWAS_list, 
         n_traits, n_scale, n_pheno_type, n_region, 
         scale_entropy, pheno_entropy, region_entropy)%>%
  mutate(# Adjust by n_trait size to find unexpectedly pleitropic and specific LD blocks
    scale_specificity = log10(n_traits) * (1 - scale_entropy),
    pheno_specificity = log10(n_traits) * (1 - pheno_entropy),
    region_specificity = log10(n_traits) * (1 - region_entropy),
    specificity_score = 0.6 * region_specificity + 0.3 * pheno_specificity + 0.1 * scale_specificity,
    pleiotropy_score =log10(n_traits) * (0.6 * region_entropy + 0.3 * pheno_entropy + 0.1 * scale_entropy))
```
```{r Figure 4.3.3B - Bar plot for Replicating Most extreme p values}
plot_df = LD_summary %>%
  pivot_longer(c(specificity_score, pleiotropy_score), names_to = "type", values_to = "score") %>%
  mutate(
    type = recode(type, specificity_score = "Specificity", pleiotropy_score  = "Pleiotropy")) %>%
  slice_max(score, n = 10, by = type)

chr_counts = plot_df %>%
  count(chr, name = "n_bar") %>%
  complete(chr = 1:22, fill = list(n_bar = 0)) %>%
  mutate(
    width = pmax(n_bar, 1),
    start = cumsum(lag(width + .5, default = 0)),
    mid = start + width / 2
  )

plot_df = plot_df %>%
  left_join(
    select(chr_counts, chr, start_chr = start),
    by = "chr"
  ) %>%
  mutate(
    rank_chr = dense_rank(locus),
    x = start_chr + rank_chr,
    label = sprintf("CHR%s:L%s (%sT)", chr, locus, n_traits),
    .by = chr
  )

x_max = max(chr_counts$start + chr_counts$width)

panel_LD = ggplot(plot_df, aes(x, score, fill = type)) +
  geom_col(width = .8) +
  geom_text(
    aes(label = label),
    angle = 90,
    hjust = -.09,
    size = 2.4
  ) +
  scale_fill_manual(
    values = c(
      Specificity = "#2166AC",
      Pleiotropy  = "#B2182B"
    )
  ) +
  scale_x_continuous(
    breaks = chr_counts$mid,
    labels = chr_counts$chr,
    limits = c(0, x_max),
    expand = expansion(mult = c(.01, .01))
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(.02, .25))
  ) +
  coord_cartesian(clip = "off") +
  labs(
    x = "Chromosome",
    y = "Pleiotropy / Specificity Score",
    fill = NULL
  ) +
  theme_classic() +
  theme(
    legend.position = "bottom",
    panel.grid = element_blank(),
    axis.text.x = element_text(face = "bold", size = 10),
    axis.title.x = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold"),
    plot.margin = margin(30, 20, 10, 10)
  )

panel_LD

ggsave(
  here("result_manuscript/F3.3B_Top_Pleiotropy_Specificity_LD_Blocks.png"),
  panel_LD,
  width = 15,
  height = 5,
  dpi = 600
)

ggsave(
  here("result_manuscript/F3.3B_Top_Pleiotropy_Specificity_LD_Blocks.pdf"),
  panel_LD,
  width = 15,
  height = 5,
  dpi = 600
)
```
# Section 4: Co-localisation, Clustering and Network Analysis for identifying latent structures
```{r Section 4 Specific library and functions}
# For Upset Graph
library(UpSetR)
library(uwot)
library(igraph)
```
```{r Figure 4.3.4A and B: UpSet hyprcoloc result Visualisation}
parse_traits_long = function(df) {
  df %>%
    mutate(
      traits = str_replace_all(traits, "SubCorVol", "SubCor_Vol"),
      traits = str_replace_all(traits, "CC_", "CC."),
      traits = str_replace_all(traits, "Mid_", "Mid."),
      traits = str_replace_all(traits, "_l\\b", ".l"),
      traits = str_replace_all(traits, "_r\\b", ".r")
    ) %>%
    mutate(row_id = row_number()) %>%
    separate_rows(traits, sep = ",") %>%
    mutate(traits = str_trim(traits)) %>%
    extract(
      traits,
      into = c("scale", "modality", "region"),
      regex = "^(global|regional|SubCor|tract)_([A-Za-z0-9]+)_?(.*)$",
      remove = FALSE) %>%
    mutate(region = na_if(region, ""))}

hyprcoloc.95cred_set = fread(here("meta_MRI_post_GWAS/supplement_results/hyprcoloc_95creditset_v2.txt"))
hyprcoloc.df <- parse_traits_long(hyprcoloc.95cred_set)

pheno_wide = hyprcoloc.df %>%
  distinct(row_id, modality) %>%
  mutate(value = 1L) %>%
  pivot_wider(
    names_from = modality,
    values_from = value,
    values_fill = 0
  ) %>%
  select(-row_id) %>%
  mutate(across(everything(), as.integer))

combo_id = apply(pheno_wide, 1, paste, collapse = "_")
combo_counts = table(combo_id)
keep_rows = as.numeric(combo_counts[combo_id]) > 2
pheno_wide_filtered = pheno_wide[keep_rows, ]


F4.3.4A = UpSetR::upset(
  as.data.frame(pheno_wide_filtered),
  order.by = "freq",
  main.bar.color = "steelblue",
  sets.bar.color = "darkorange",
  text.scale = 1.3,
  mainbar.y.label = "Colocalising trait combinations",
  sets.x.label = "Loci per phenotype")


# For regions
region_membership <- hyprcoloc.df %>%
  filter(!is.na(region)) %>%
  distinct(row_id, region)

region_wide = region_membership %>%
  mutate(value = 1L) %>%
  pivot_wider(
    names_from = region,
    values_from = value,
    values_fill = 0
  ) %>%
  select(-row_id) %>%
  mutate(across(everything(), as.integer))

# Keep combinations occurring > 2 times
combo_id = apply(region_wide, 1, paste, collapse = "_")
combo_counts = table(combo_id)
region_wide_filtered = region_wide[as.numeric(combo_counts[combo_id]) > 2,]

F4.3.4B = UpSetR::upset(as.data.frame(region_wide_filtered),
  order.by = "freq",
  main.bar.color = "steelblue",
  sets.bar.color = "darkorange",
  mainbar.y.label = "Colocalising region combinations",
  sets.x.label = "Loci per region")


```
Generation of genetic correlation Matrix as shown in MRI_within_rg_matrix_construction.Rmd
```{r igraph network construction}
# read in data, df formatting and clean-up
rg_network.df = fread(here("script_manuscript/network_clustering_component/rg_matrix.QCed.formatted.txt")) %>% 
  select(p1_base, p2_base, rg, p) %>% 
  mutate(rg = pmax(pmin(rg, 1), -1), p_fdr = p.adjust(p, method = "fdr")) # some rg > 1 correct them to 1

rg_network.df = sup_table_name_cleaner(rg_network.df, c("p1_base", "p2_base"))

# define edge and node-associated properties

## Edge: rg and distance
rg_edge = rg_network.df %>% filter(p_fdr < 0.05) %>% # filter out weak fdr p values
  select(p1_base, p2_base, rg) %>% 
  rename(from = p1_base, to = p2_base)

### add between region distance
geo_distance.df = as.matrix(fread(here("script_manuscript/network_clustering_component/geodesic_distance_HCP.csv"), header = TRUE))[1:180,1:180]
geo_distance.colnames = fread(here("script_manuscript/network_clustering_component/geodesic_distance_HCP_colnames.csv"))
colnames(geo_distance.df) = geo_distance.colnames$region
rownames(geo_distance.df) = geo_distance.colnames$region

rg_edge[, from_region := sub("^regional_[^_]+_", "", from)]
rg_edge[, to_region   := sub("^regional_[^_]+_", "", to)]
rg_edge[, geo_distance := NA_real_]
idx = grepl("^regional_", rg_edge$from) & grepl("^regional_", rg_edge$to)
rg_edge[idx, geo_distance := geo_distance.df[cbind(from_region, to_region)]]

rg_edge = rg_edge %>% mutate(weight = (rg + 1) / 2) %>% select(from, to, rg, weight, geo_distance)

## Node: MRI categories and h2
rg_node = fread(here("result_manuscript/ST3_UKB_and_meta_h2.txt")) %>% 
  mutate(name = paste(MRI_scale, MRI_pheno, region, sep = "_"), name = gsub("_NA", "", name)) %>% 
  select(name, MRI_scale, MRI_pheno, region, UKB_h2)

# network construction
rg_network = graph_from_data_frame(d = rg_edge, vertices = rg_node, directed = FALSE)
saveRDS(rg_network, here("script_manuscript/network_clustering_component/rg_network.Rds"))
```
```{r leiden clustering and investigation on key network properties}
# leiden clustering
resolutions = seq(0.1, 2, by = 0.05)

leiden.opt = sapply(resolutions, function(res) {
  cl = cluster_leiden(rg_network, weights = E(rg_network)$rg, objective_function = "modularity", resolution_parameter = res)
  modularity(rg_network, membership(cl))})

best_res = resolutions[which.max(leiden.opt)]

leiden_cl = cluster_leiden(rg_network, weights = E(rg_network)$rg, objective_function = "modularity", resolution = 0.5) # this was determined step-wise gridsearch
sizes(leiden_cl)
leiden_membership.df = tibble(
  Trait = names(membership(leiden_cl)),
  ld_cl = factor(membership(leiden_cl)))

# Hierarchical Clustering
rg_matrix = as.matrix(as_adjacency_matrix(rg_network, attr = "weight", sparse = FALSE)) # retrieve transformed rg [-1,1] to [0,1] as matrix
hc = hclust(as.dist(1 -rg_matrix), method = "average") # Change adjacency matrix to distance, average linked
hc_cl = dynamicTreeCut::cutreeDynamic(dendro = hc, distM = as.matrix(as.dist(1 -rg_matrix)), minClusterSize = 233.2) 
hc_membership.df = data.frame(Trait = colnames(rg_matrix), hc_cl = hc_cl)

# Note on minClusterSize: Force big clusters, needs to include at least 10% of the total nodes per cluster (2332*0.1). Other no clusters will be formed.
```
```{r UMAP Visualisation}
umap = umap2(1 - rg_matrix, n_neighbors = 50, min_dist = 0.05) # UMAP

# UMAP data table construction
umap.df = as.data.frame(umap) %>% 
  rename(x = V1, y = V2) %>% 
  mutate(GWAS = rownames(.), GWAS_name = rownames(.)) %>% 
  sup_table_formatter()

umap.df = umap.df %>% full_join(leiden_membership.df, join_by("GWAS" == "Trait")) %>% full_join(hc_membership.df, join_by("GWAS" == "Trait"))

umap.df = umap.df %>% 
  mutate(
    SubCor_group = case_when(
      !str_detect(GWAS, "^SubCor_Vol") ~ NA_character_,
      str_detect(GWAS, "^SubCor_Vol_CC.") ~ "Corpus Callosum",
      str_detect(GWAS, "Ventricle|Vent$") ~ "Ventricles",
      str_detect(GWAS, regex("Caudate|Accumbens|Putamen|Pallidum|Thalamus|Amygdala|Hippocampus|VentralDC", ignore_case = TRUE)) ~ "Deep Grey Matter",
      str_detect(GWAS,regex("Brain.Stem|Cerebellum", ignore_case = TRUE)) ~ "Hindbrain", TRUE ~ "Other"))

# For MRI_Scale
UMAP.scale = umap.df %>%
  {
    ggplot() +
      geom_point( # make sure global dots plotted last and visable
        data = subset(., MRI_scale != "global"),
        aes(x, y, colour = MRI_scale),
        size = 1.5, alpha = 0.8
      ) +
      geom_point(
        data = subset(., MRI_scale == "global"),
        aes(x, y, colour = MRI_scale),
        size = 1.5, alpha = 0.8
      )} +
  scale_color_manual(
    values = c(
      regional = "#FDBF6F",
      global   = "#E31A1C",
      tract    = "#CAB2D6",
      SubCor   = "grey30"),
    labels = c(
      global   = "Glb",
      regional = "Rgnl",
      SubCor   = "SubB.Vol",
      tract    = "WM.Tracts")) +
  labs(
    colour = "Region Types") +
  theme_pubr() +
  theme(legend.position = "bottom", legend.box = "vertical")

# For MRI_pheno
UMAP.pheno = umap.df %>% ggplot(aes(x,y,colour = MRI_pheno)) +
  geom_point(size = 1.5, alpha = 0.8, position = position_jitter(width = 0.2, height = 0.2)) + 
  scale_colour_manual(values = MRI_pheno_all.colors) +
  theme_pubr() +
   labs(colour = "MRI Modality")+
  theme(legend.position = "bottom")

# For ld_cl
UMAP.leiden = umap.df %>% ggplot(aes(x,y,colour = ld_cl)) +
  geom_point(size = 1.5, alpha = 0.8) + 
  theme_pubr() +
   labs(colour = "Leiden Clustering")+
  theme(legend.position = "bottom")

# For hc_cl
UMAP.hc = umap.df %>% ggplot(aes(x,y,colour = as.character(hc_cl))) +
  geom_point(size = 1.5, alpha = 0.8) + 
  theme_pubr() +
   labs(colour = "Hiearchical Clustering")+
  theme(legend.position = "bottom")
```
```{r F4.3.5 Construction}
# Individual graph tweaks 
common_theme = theme(legend.position = "bottom", legend.box = "vertical", legend.title = element_text(size = 9), 
                     legend.text = element_text(size = 8), legend.key.size = unit(0.3, "cm"), axis.title = element_blank())

UMAP.scale = UMAP.scale + common_theme + guides(colour = guide_legend(order = 1))
UMAP.pheno = UMAP.pheno + common_theme + guides(colour = guide_legend(order = 3))
UMAP.hc = UMAP.hc + common_theme + guides(colour = guide_legend(order = 4))
UMAP.leiden = UMAP.leiden + common_theme + guides(colour = guide_legend(order = 5))


F4.3.5 = (UMAP.scale + UMAP.pheno + UMAP.hc + UMAP.leiden) + plot_layout(guides = "collect") + plot_annotation(tag_levels = "A")

ggsave(
  here("result_manuscript/F4.3.5A-D.png"),
  plot = F4.3.5,
  width = 11.8,
  height = 5.6,
  units = "in",
  dpi = 600)

ggsave(
  here("result_manuscript/F4.3.5A-D.pdf"),
  plot = F4.3.5,
  width = 11.8,
  height = 5.6,
  units = "in",
  dpi = 600)

```
```{r Statistical Analysis for Cortical Region: Effect of Geodistance and Clustering membership on rg}
# include Leiden and Hierarchical clustering information
rg.stat = rg_edge %>% 
  left_join(umap.df %>% select(GWAS, ld_cl, hc_cl) %>% 
              rename(from_ld = ld_cl, from_hc = hc_cl),by = c("from" = "GWAS")) %>%
  left_join(umap.df %>%
              select(GWAS, ld_cl, hc_cl) %>%
              rename(to_ld = ld_cl, to_hc = hc_cl), by = c("to" = "GWAS")) %>%
  mutate(same_ld = from_ld == to_ld, same_hc = from_hc == to_hc, 
         ld_pair = paste(pmin(as.numeric(from_ld), as.numeric(to_ld)), 
                         pmax(as.numeric(from_ld), as.numeric(to_ld)), sep = "-")) # find same cluster pair and different cluster pairs

# average rg with same clustering
rg.same = rg.stat %>%
  group_by(same_ld) %>%
  summarise(
    mean_rg = mean(rg, na.rm = TRUE),
    mean_abs_rg = mean(abs(rg), na.rm = TRUE)
  )

# average rg across all cluster pairing possible
rg.pairs = rg.stat %>%
  group_by(ld_pair) %>%
  summarise(
    mean_rg = mean(rg),
    mean_abs_rg = mean(abs(rg))
  ) %>%
  arrange(desc(mean_abs_rg))

# rg with geo_distance only
rg_model.geo_dist = lm(rg ~ geo_distance , data = rg.stat)
summary(rg_model.geo_dist)

# absolute rg with geo_distance only
rg_model.geo_dist.abs = lm(abs(rg) ~ geo_distance, data = rg.stat)
summary(rg_model.geo_dist.abs)

# absolute rg with geo_distance and same_ld [0/1]
rg_model.geo_dist_same_ld.abs = lm(abs(rg) ~ geo_distance + same_ld, data = rg.stat)
summary(rg_model.geo_dist_same_ld.abs)

# absolute rg with ld pair considerations
rg_model.geo_dist_ld_pair.abs = lm(abs(rg) ~ geo_distance + ld_pair, data = rg.stat)
summary(rg_model.geo_dist_ld_pair.abs)

anova(rg_model.geo_dist.abs, rg_model.geo_dist_same_ld.abs, rg_model.geo_dist_ld_pair.abs)
```
# Section 5: FLAMES and Enrichment Analysis
```{r Section 5 specific library}
library(circlize)
library(AnnotationDbi)
library(org.Hs.eg.db)

# find overlap for chord graphs
find_overlap_matrix = function(df, group_var){
  genes_by_group = df %>%
    group_by(.data[[group_var]]) %>%
    summarise(genes = list(unique(symbol)), .groups = "drop")

  groups = genes_by_group[[group_var]]

  mat = matrix(0, nrow = length(groups), ncol = length(groups), dimnames = list(groups, groups))

  for(i in seq_along(groups)){
    for(j in seq_along(groups)){
      gi = genes_by_group$genes[[i]]
      gj = genes_by_group$genes[[j]]
      mat[i,j] <- length(intersect(gi, gj))
    }
  }

  diag(mat) = 0
  mat}

MRI_family.colours = c(
  "Macrostructural Morphology" = "#E64B35",
  "White Matter Integrity"     = "#4DBBD5",
  "Cortical Geometry"          = "#00A087",
  "Water Diffusivity"          = "#3C5488",
  "Cortical Thickness"         = "#F39B7F",
  "Orientation Dispersion"     = "#8491B4")

plot_chord = function(overlap.mat, grid.col = NULL){
  # reset if plotted chord beforehand
  circos.clear()
  
  circos.par(start.degree = 90, gap.after = rep(8, nrow(overlap.mat)), track.margin = c(0.01, 0.01))
  
  chordDiagram(overlap.mat, grid.col = grid.col, transparency = 0.4, annotationTrack = "grid", preAllocateTracks = 1)
  
  circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y){
    sector = get.cell.meta.data("sector.index")
    xlim = get.cell.meta.data("xlim")
    ylim = get.cell.meta.data("ylim")
    
    circos.text(mean(xlim), ylim[2], sector, 
                facing = "bending.outside", niceFacing = TRUE, 
                cex = 0.8, font = 2)}, 
    bg.border = NA)
  
  circos.clear()
}

```
## FLAMES
FLAMES genes were prioritised through FLAMES_README.Rmd all bash scripts with "FLAMES_" prefix
```{r write SF6: FLAMES genes, eval=FALSE, include=FALSE}
FLAMES = list.files(here("meta_MRI_post_GWAS/UKB_FLAMES"), recursive = TRUE, pattern = "FLAMES_scores.pred", full.names = TRUE)

FLAMES.pred.df = map_dfr(FLAMES, function(file) {
  df = fread(file)
  if (nrow(df) == 0) {
    message("Skipping due to <75% cumulative precision: ", basename(dirname(file)))
    return(NULL)}
  df %>% mutate(source_file = basename(dirname(file)))})

# Save supplement results:
FLAMES.score.pred.df.formatted = FLAMES.pred.df %>% 
  rename(phenotype = source_file) %>% 
  select(phenotype, symbol, ensg, FLAMES_scaled, estimated_cumulative_precision) %>% 
  sup_table_name_cleaner(col_name = "phenotype") %>% 
  mutate(GWAS_name = phenotype) %>%
  sup_table_formatter() %>%
  select(phenotype, MRI_scale, MRI_pheno, region, symbol, ensg, FLAMES_scaled, estimated_cumulative_precision)

write.table(FLAMES.score.pred.df.formatted, 
            here("result_manuscript/SF8_FLAMES_gene_prioritisation_results.txt"),
            quote = FALSE, row.names = FALSE)
```
```{r F4.3.6 Chord Graph on overlapping genes}
# read and formatt FLAMES gene dataset
FLAMES.score.pred.df.formatted = fread(here("result_manuscript/SF8_FLAMES_gene_prioritisation_results.txt"))
FLAMES.unique = FLAMES.score.pred.df.formatted %>%
  distinct(phenotype, MRI_scale, MRI_pheno, region, symbol)

FLAMES.unique = FLAMES.unique %>%
  mutate(
    family = case_when(
      MRI_pheno %in% c("SA","GMV","FI","IC") ~ "Macrostructural Morphology",
      MRI_pheno %in% c("FA","ICVF") ~ "White Matter Integrity",
      MRI_pheno %in% c("GC","MC") ~ "Cortical Geometry",
      MRI_pheno %in% c("MD","ISOVF") ~ "Water Diffusivity",
      MRI_pheno == "CT" ~ "Cortical Thickness",
      MRI_pheno == "OD" ~ "Orientation Dispersion",
      MRI_pheno == "Vol" &
        region %in% c("choroid.plexus", "CSF", "Optic.Chiasm",
                      "3rd.Ventricle", "4th.Ventricle", "Inf.Lat.Vent") ~ "Water Diffusivity",
      MRI_pheno == "Vol" &
        !region %in% c("choroid.plexus", "CSF", "Optic.Chiasm", 
                       "3rd.Ventricle", "4th.Ventricle", "Inf.Lat.Vent") ~ "Macrostructural Morphology",
      TRUE ~ NA_character_))

# build overlap matrix
scale.overlap_mat = find_overlap_matrix(FLAMES.unique, "MRI_scale")
pheno.overlap_mat = find_overlap_matrix(FLAMES.unique, "MRI_pheno")
family.overlap_mat = find_overlap_matrix(FLAMES.unique, "family")

# chord plotting
plot_chord(scale.overlap_mat)
plot_chord(pheno.overlap_mat, MRI_pheno_all.colors)
plot_chord(family.overlap_mat, MRI_family.colours)
```
## Cell Type+Timepoint+Location Enrichment Analysis
```{r Pseudoblock Matrix Generation: WANG, eval=FALSE, include=FALSE}
FLAMES.family = FLAMES.unique %>% distinct(symbol, family)

Wang_human = list.files(here("script_manuscript/enrichment_component/celltype"), pattern = "Wang2024_Human", full.names = TRUE) 
Wang_human_level2 = Wang_human[grepl("level2", basename(Wang_human))]

# Generate pseudo-blocks
wang_long = rbindlist(
  lapply(Wang_human_level2, function(f){

    dt = fread(f)
    avg_dt = dt[, .(GENE, avg_expr = Average)] # find average expression value per file
    
    # remove Average column
    dt[, Average := NULL]

    # remove prefix/suffix
    fname = basename(f)
    meta = gsub("^\\d+_", "", fname)
    meta = gsub("^Wang2024_Human_2024_", "", meta)
    meta = gsub("_level2\\.txt$", "", meta)
    
    # Find brain part and developmental timepoint
    parts = strsplit(meta, "_")[[1]]
    timepoint = parts[1]
    region = paste(parts[-1], collapse = "_")

    long = melt(dt, id.vars = "GENE", variable.name = "celltype", value.name = "expr")
    long = merge(long, avg_dt, by = "GENE")
    long[, bin := paste(celltype, timepoint, region, sep = "_")]

    long[, .(GENE, celltype, timepoint, region, bin, expr, avg_expr)]}), use.names = TRUE)

# construct covars for glm modeling
gene_covars = data.frame(GENE = unique(wang_long$GENE))

# add gene symbol to ENSEMBL
gene_map = AnnotationDbi::select(org.Hs.eg.db, keys = unique(wang_long$GENE), keytype = "ENSEMBL", columns = "SYMBOL")
gene_covars = left_join(gene_covars, gene_map, by = c("GENE" = "ENSEMBL"))

# add gene size
ENSG.gene.loc = fread(glue("{path_vw.toolbox}magma/ENSG.gene.loc"))
gene_size = ENSG.gene.loc %>% transmute(ENSEMBL = V1, Symbol = V5, gene_length = V4 - V3 + 1)

gene_covars = gene_covars %>% 
  left_join(gene_size, by = c("GENE" = "ENSEMBL")) %>%
  mutate(log_length = log10(gene_length))

# filter for gene symbol version that has most FLAMES gene overlap 
for(fam in unique(FLAMES.family$family)){

  fam_genes = FLAMES.family %>%
    filter(family == fam) %>%
    pull(symbol)
  
  # use Symbol from magma definition as they have more overlap with FLAMES gene
  gene_covars[[paste0("is_", fam)]] = as.integer(gene_covars$Symbol %in% fam_genes)}

colnames(gene_covars) = gsub(" ", "_", colnames(gene_covars))

write.table(gene_covars, 
            here("script_manuscript/enrichment_component/gene_covar_annotated.txt"),
            quote = FALSE, row.names = FALSE)

write.table(wang_long, 
            here("script_manuscript/enrichment_component/psuedobulk_wang.txt"),
            quote = FALSE, row.names = FALSE)
```
```{r Logistic Regression Modelling}
wang_long = fread(here("script_manuscript/enrichment_component/psuedobulk_wang.txt"))
gene_covars = fread(here("script_manuscript/enrichment_component/gene_covar_annotated.txt"))

# set up for loop
results = list()
families = colnames(gene_covars)[grepl("is_", colnames(gene_covars))]

for(fam in families){
  for(current_bin in unique(wang_long$bin)){

    bin_df = wang_long %>% 
      filter(bin == current_bin) %>%
      select(GENE, expr, avg_expr)

    mdf = gene_covars %>% 
      left_join(bin_df %>% select(GENE, expr, avg_expr), by = "GENE") %>%
      mutate(rel_expr = expr - avg_expr) %>%
      filter(
        !is.na(log_length),
        !is.na(expr),
        !is.na(avg_expr))

    fit = glm(mdf[[fam]] ~ rel_expr + log_length, family = binomial(), data = mdf)

    cf = summary(fit)$coefficients

    beta = cf["rel_expr", "Estimate"]
    se   = cf["rel_expr", "Std. Error"]

    z = beta / se
    p = pnorm(z, lower.tail = FALSE) # only finds enrichment not depletion

    results[[length(results) + 1]] =
      data.frame(
        family = fam,
        bin = current_bin,
        beta = beta,
        se = se,
        p = p)}}

# reduce and format celltype result tables
FLAMES.cell = bind_rows(results)

FLAMES.cell = FLAMES.cell %>%
  group_by(bin) %>%
  mutate(p.fdr.bin = p.adjust(p, "fdr")) %>%
  ungroup() %>%
  group_by(family) %>%
  mutate(p.fdr.family = p.adjust(p, "fdr")) %>%
  ungroup() %>%
  mutate(p.fdr = p.adjust(p, "fdr"))

sig.FLAMES.cell = FLAMES.cell %>% filter(p < 0.05)

regions = unique(wang_long$region)
timepoints = unique(wang_long$timepoint)

sig.FLAMES.cell = sig.FLAMES.cell %>%
  mutate(
    region = str_extract(bin, paste(regions, collapse = "|")),
    timepoint = str_extract(bin, paste(timepoints, collapse = "|")),
    celltype = bin %>% str_remove(paste0("_", timepoint, "_", region)))
```
```{r F4.3.7 Concentric circle plot for Cell type enrichment stratified by developmental stage and brain location}
circle_sizes = c(FirstTrimester = 2, SecondTrimester = 4, ThirdTrimester = 6, Infancy = 8, Adolescence = 10)
circle_colours <- c(
  FirstTrimester = "#d73027",
  SecondTrimester = "#fc8d59",
  ThirdTrimester = "#fee090",
  Infancy = "#91bfdb",
  Adolescence = "#4575b4")

region_order = c("Forebrain", "Telencephalon", "Neocortex", 
                 "PrefrontalCortex", "VisualCortex", 
                 "BrodmannArea9", "BrodmannArea10", "BrodmannArea17")

celltype_order = c("Astrocyte", "Oligodendrocyte",
                   "Cajal Retzius Cell", 
                   "IPC EN", "IPC Glia", "Radial Glia")

timepoint_order = c("FirstTrimester", "SecondTrimester", "ThirdTrimester", "Infancy", "Adolescence")

FLAMES.celltype.plot = sig.FLAMES.cell %>% filter(p.fdr < 0.05) %>%
  select(family, celltype, timepoint, region) %>% 
  filter(!celltype %in% c("Unknown", "Vascular")) %>%
  mutate(
    family = gsub("is_", "", family),
    family = gsub("_", " ", family),
    celltype = gsub("_", " ", celltype),
    celltype = gsub("glia", "Glia", celltype),
    celltype = gsub("cell", "Cell", celltype),
    circle_sizes = circle_sizes[timepoint],
    circle_colours = circle_colours[timepoint],
    timepoint = factor(timepoint, levels = timepoint_order),
    region = factor(region, levels = region_order),
    celltype = factor(celltype, levels = celltype_order)) %>%
  arrange(desc(circle_sizes))

# Circle plot
F4.3.7 = ggplot(FLAMES.celltype.plot, aes(x = region, y = celltype)) +
  geom_point(aes(size = timepoint, colour = timepoint),
             shape = 1, stroke = 1.2) +
  scale_size_manual(values = circle_sizes, name = "Developmental stage") +
  scale_colour_manual(values = circle_colours, name = "Developmental stage") +
  facet_wrap(~ family) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, colour = "black"),
    axis.text.y = element_text(colour = "black"),
    strip.background = element_blank(),
    panel.grid.major.x = element_line(colour = "grey90",linewidth = 0.2),
    panel.grid.major.y = element_line(colour = "grey90", linewidth = 0.2),
    panel.grid.minor = element_blank()) +
  labs(x = NULL, y = NULL)


ggsave(
  here("result_manuscript/F4.3.7.png"),
  plot = F4.3.7,
  width = 9,
  height = 6,
  units = "in",
  dpi = 600)

ggsave(
  here("result_manuscript/F4.3.7.pdf"),
  plot = F4.3.7,
  width = 9,
  height = 6,
  units = "in",
  dpi = 600)
```
```{r write ST7 sig.CellType Result}
ST7_sig.celltype_Wang2025 = sig.FLAMES.cell %>% filter(p.fdr < 0.05) %>%
  filter(!celltype %in% c("Unknown", "Vascular")) %>%
  mutate(
    family = gsub("is_", "", family),
    family = gsub("_", " ", family),
    celltype = gsub("_", " ", celltype),
    celltype = gsub("glia", "Glia", celltype),
    celltype = gsub("cell", "Cell", celltype))

write.table(ST7_sig.celltype_Wang2025, here("result_manuscript/ST7_sig.celltype_Wang2025.txt"), quote = F, row.names = F)
```

## Gene-set Functional Analysis
Done locally through Clusterprofiler.Rmd

Done locally through Clusterprofiler, this is due to hpc environment do not have the required modules loaded. For scripts check ClusterProfiler_Enrichment.Rmd

# Section 6: Autism gene enrichment
```{r SFARI gene list and custom term2gene set up for MRI modality family}
SFARI_gene.list = fread(here("script_manuscript/enrichment_component/SFARI-Gene_genes_07-12-2026release_09-10-2026export.csv"))

# syndromic[0/1] = Known Syndrome with reported ASD gene, different from gene category syndromic which means other symptoms were associated with it (ID)
# 1,2,3 = Strength of Confirmed ASD reports
# Eagle = Higher means more ASD-specific than NDD-specific

SFARI_gene.ASD_syndromic = SFARI_gene.list %>% filter(syndromic == 1, `gene-score` %in% c(1, 2)) # ASD Genes known for other syndromes
SFARI_gene.ASD_confidence = SFARI_gene.list %>% filter(syndromic == 0, `gene-score` %in% c(1, 2)) # High-confidence ASD genes
SFARI_gene.ASD_specific = SFARI_gene.list %>% filter(eagle >= 12) # ASD-specific genes
SFARI_gene.ASD_specific.nosyndrome = SFARI_gene.list %>% filter(syndromic == 0, eagle >= 12) # non-syndromic ASD-specific genes

SFARI_gene.list.cat = list(
  All = SFARI_gene.list, Syndromic = SFARI_gene.ASD_syndromic, HighConfidence = SFARI_gene.ASD_confidence,
  HighEagle = SFARI_gene.ASD_specific, HighEagle_NoSyndrome = SFARI_gene.ASD_specific.nosyndrome)

# Build custom gene-set for over-representation test
FLAMES_gene_results = fread(here("result_manuscript/SF8_FLAMES_gene_prioritisation_results.txt"))

FLAMES_gene_results.f = FLAMES_gene_results %>%
  mutate(
    family = case_when(
      MRI_pheno %in% c("SA","GMV","FI","IC") ~ "Macrostructural Morphology",
      MRI_pheno %in% c("FA","ICVF") ~ "White Matter Integrity",
      MRI_pheno %in% c("GC","MC") ~ "Cortical Geometry",
      MRI_pheno %in% c("MD","ISOVF") ~ "Water Diffusivity",
      MRI_pheno == "CT" ~ "Cortical Thickness",
      MRI_pheno == "OD" ~ "Orientation Dispersion",
      MRI_pheno == "Vol" &
        region %in% c("choroid.plexus", "CSF", "Optic.Chiasm",
                      "3rd.Ventricle", "4th.Ventricle", "Inf.Lat.Vent") ~ "Water Diffusivity",
      MRI_pheno == "Vol" &
        !region %in% c("choroid.plexus", "CSF", "Optic.Chiasm", 
                       "3rd.Ventricle", "4th.Ventricle", "Inf.Lat.Vent") ~ "Macrostructural Morphology",
      TRUE ~ NA_character_))

MRI_modality.family2gene = FLAMES_gene_results.f %>%
  select(family, ensg) %>%
  distinct()

MRI_modality.family2gene.all = MRI_modality.family2gene %>% mutate(family = "MRI") %>% distinct()

# background genes
FLAMES_background = fread(glue("{path_vw.toolbox}FLAMES/16_Feb_2021_gene_loc.txt"))
```
```{r ORA:fisher test function}
gene.ORA = function(sfari_genes, family2gene, universe){
  N = length(universe)
  K = sum(universe %in% sfari_genes)
  
  family2gene %>% distinct(family, ensg) %>% 
    group_by(family) %>%
    summarise(
      family_size = n(),
      overlap = sum(ensg %in% sfari_genes),
      overlap_genes = paste(ensg[ensg %in% sfari_genes],collapse = ";"),
      
      GeneRatio = paste0(overlap, "/", K),
      BgRatio = paste0(family_size, "/", N),
      FoldEnrichment = (overlap / K) / (family_size / N),
  
      p = fisher.test(matrix(c(overlap, K - overlap, family_size - overlap, N - K - family_size + overlap), nrow = 2),
                      alternative = "greater")$p.value,
      OR = fisher.test(matrix(c(overlap, K - overlap, family_size - overlap, N - K - family_size + overlap), nrow = 2),
                       alternative = "greater")$estimate, .groups = "drop") %>%
    mutate(p.fdr = p.adjust(p, method = "BH")) %>%
    arrange(p.fdr)}
```
```{r ORA results}
# Run ORA
SFARI_ORA = map_dfr(names(SFARI_gene.list.cat), \(g){
  gene.ORA(sfari_genes = unique(SFARI_gene.list.cat[[g]]$`ensembl-id`),
           family2gene = MRI_modality.family2gene,
           universe = unique(FLAMES_background$ENSGID)) %>%
    mutate(SFARI_group = g)})

SFARI_ORA_allMRI = map_dfr(names(SFARI_gene.list.cat), \(g){
  gene.ORA(
    sfari_genes = unique(SFARI_gene.list.cat[[g]]$`ensembl-id`),
    family2gene = MRI_modality.family2gene.all,
    universe = unique(FLAMES_background$ENSGID)) %>%
    mutate(SFARI_group = g)})

# combine table
SFARI_ORA.results = full_join(SFARI_ORA, SFARI_ORA_allMRI)
sig.SFARI_ORA.results = SFARI_ORA.results %>% filter(p.fdr < 0.05)
```
```{r write ST8 SFARI_ORA.results}
write.table(SFARI_ORA.results, 
            here("result_manuscript/ST8_Autism_vs_MRI_family_ORA.txt"),
            quote = FALSE, row.names = FALSE)
```

# Section 7: rg with Autism and Autistic Traits
```{r section 7 specific packages}
# brain visualisation
library(ggseg)
library(ggsegGlasser) # cortical

# MR
library(TwoSampleMR)

# network
library(igraph)
```

## rg visualisation
```{r Autistic traits and Autism vs MRI GWAS, eval=FALSE, include=FALSE}
# AQ
AQ.path = list.files(glue("{path_UKB.vw}SPQ/SPQ_post_GWAS/ldsc_h2_and_rg"), pattern ="rg_AQ_vs_.*\\.log$", full.names = TRUE)
AQ_global.path = list.files(here("script_manuscript/genetic_correlation_component/rg_result/"), pattern ="UKB_AQ_.*\\.log$", full.names = TRUE)
AQ.path_all = c(AQ.path, AQ_global.path)

AQ.rg = ldsc_rg_batch_read(AQ.path_all)  %>% 
  mutate(p.exp_adj = p.adjust(p, method = "fdr"), 
         p.exp_label = case_when(p.exp_adj <0.001 ~ "***", p.exp_adj <0.01 ~ "**", p.exp_adj <0.05 ~ "*", TRUE ~ NA_character_)) %>% 
  rename(GWAS_name = p2_base) %>% mutate(GWAS_name = gsub("GWAS_", "", GWAS_name))

AQ.rg.formatted = sup_table_name_cleaner(AQ.rg) %>%
  sup_table_formatter() %>%
  group_by(MRI_scale, MRI_pheno) %>%
  mutate(
    p.idp_adj = p.adjust(p, method = "fdr"),
    p.idp_label = case_when(
      p.idp_adj < 0.001 ~ "***",
      p.idp_adj < 0.01  ~ "**",
      p.idp_adj < 0.05  ~ "*",
      TRUE ~ NA_character_)) %>%
  ungroup() %>%
  mutate(p1_base = "AQ_total_score") %>%
  relocate(p, p.label, p.idp_adj, p.idp_label, p.exp_adj, p.exp_label, .after = last_col())

# SPQ
SPQ.path = list.files(glue("{path_UKB.vw}SPQ/SPQ_post_GWAS/ldsc_h2_and_rg"), pattern ="rg_SPQ_vs_.*\\.log$", full.names = TRUE)
SPQ_global.path = list.files(here("script_manuscript/genetic_correlation_component/rg_result/"), pattern ="total_vs_.*\\.log$", full.names = TRUE)
SPQ.path_all = c(SPQ.path, SPQ_global.path)
SPQ.path_all = SPQ.path_all[!grepl("rg_SPQ_vs_autismQ|rg_SPQ_vs_PSYCH", SPQ.path_all)]

SPQ.rg = ldsc_rg_batch_read(SPQ.path_all)  %>% 
  mutate(p.exp_adj = p.adjust(p, method = "fdr"), 
         p.exp_label = case_when(p.exp_adj <0.001 ~ "***", p.exp_adj <0.01 ~ "**", p.exp_adj <0.05 ~ "*", TRUE ~ NA_character_)) %>% 
  rename(GWAS_name = p2_base)

SPQ.rg.formatted = sup_table_name_cleaner(SPQ.rg) %>%
  sup_table_formatter() %>%
  group_by(MRI_scale, MRI_pheno) %>%
  mutate(
    p.idp_adj = p.adjust(p, method = "fdr"),
    p.idp_label = case_when(
      p.idp_adj < 0.001 ~ "***",
      p.idp_adj < 0.01  ~ "**",
      p.idp_adj < 0.05  ~ "*",
      TRUE ~ NA_character_)) %>%
  ungroup() %>%
  mutate(p1_base = "SPQ_total_score") %>%
  relocate(p, p.label, p.idp_adj, p.idp_label, p.exp_adj, p.exp_label, .after = last_col())

# Autism
ASD.path = list.files(glue("{path_UKB.vw}SPQ/SPQ_post_GWAS/ldsc_h2_and_rg"), pattern ="rg_ASD_vs_.*\\.log$", full.names = TRUE)
ASD_global.path = list.files(here("script_manuscript/genetic_correlation_component/rg_result/"), pattern ="Autism_Grove_.*\\.log$", full.names = TRUE)
ASD.path_all = c(ASD.path, ASD_global.path)

ASD.rg = ldsc_rg_batch_read(ASD.path_all)  %>% 
  mutate(p.exp_adj = p.adjust(p, method = "fdr"), 
         p.exp_label = case_when(p.exp_adj <0.001 ~ "***", p.exp_adj <0.01 ~ "**", p.exp_adj <0.05 ~ "*", TRUE ~ NA_character_)) %>% 
  rename(GWAS_name = p2_base)

ASD.rg.formatted = sup_table_name_cleaner(ASD.rg) %>%
  sup_table_formatter() %>%
  group_by(MRI_scale, MRI_pheno) %>%
  mutate(
    p.idp_adj = p.adjust(p, method = "fdr"),
    p.idp_label = case_when(
      p.idp_adj < 0.001 ~ "***",
      p.idp_adj < 0.01  ~ "**",
      p.idp_adj < 0.05  ~ "*",
      TRUE ~ NA_character_)) %>%
  ungroup() %>%
  mutate(p1_base = "Autism_Grove") %>%
  relocate(p, p.label, p.idp_adj, p.idp_label, p.exp_adj, p.exp_label, .after = last_col())

# Autism: Daner
ASD_Daner.path = list.files(here("script_manuscript/genetic_correlation_component/rg_result/"), pattern ="ASDraw_.*\\.log$", full.names = TRUE)

ASD_Daner.rg = ldsc_rg_batch_read(ASD_Daner.path)  %>% 
  mutate(p.exp_adj = p.adjust(p, method = "fdr"), 
         p.exp_label = case_when(p.exp_adj <0.001 ~ "***", p.exp_adj <0.01 ~ "**", p.exp_adj <0.05 ~ "*", TRUE ~ NA_character_)) %>% 
  rename(GWAS_name = p2_base)

ASD_Daner.rg.formatted = sup_table_name_cleaner(ASD_Daner.rg) %>%
  sup_table_formatter() %>%
  group_by(MRI_scale, MRI_pheno) %>%
  mutate(
    p.idp_adj = p.adjust(p, method = "fdr"),
    p.idp_label = case_when(
      p.idp_adj < 0.001 ~ "***",
      p.idp_adj < 0.01  ~ "**",
      p.idp_adj < 0.05  ~ "*",
      TRUE ~ NA_character_)) %>%
  ungroup() %>%
  mutate(p1_base = "Autism_Daner") %>%
  relocate(p, p.label, p.idp_adj, p.idp_label, p.exp_adj, p.exp_label, .after = last_col())

# Autism Stratified by Age: Early
ASD.AaD_early.path = list.files(here("script_manuscript/genetic_correlation_component/rg_result/"), pattern ="up_early_.*\\.log", full.names = TRUE)
ASD.AaD_early.rg = ldsc_rg_batch_read(ASD.AaD_early.path)  %>% 
  mutate(p.exp_adj = p.adjust(p, method = "fdr"), 
         p.exp_label = case_when(p.exp_adj <0.001 ~ "***", p.exp_adj <0.01 ~ "**", p.exp_adj <0.05 ~ "*", TRUE ~ NA_character_)) %>% 
  rename(GWAS_name = p2_base)

ASD.AaD_early.rg.formatted = sup_table_name_cleaner(ASD.AaD_early.rg) %>%
  sup_table_formatter() %>%
  group_by(MRI_scale, MRI_pheno) %>% # macro-idp = 181; micro-idp = 208; subcor_vol = 26
  mutate(
    p.idp_adj = p.adjust(p, method = "fdr"),
    p.idp_label = case_when(
      p.idp_adj < 0.001 ~ "***",
      p.idp_adj < 0.01  ~ "**",
      p.idp_adj < 0.05  ~ "*",
      TRUE ~ NA_character_)) %>%
  ungroup() %>%
  mutate(p1_base = "Autism_Early_Diagnosed") %>%
  relocate(p, p.label, p.idp_adj, p.idp_label, p.exp_adj, p.exp_label, .after = last_col())

# Autism Stratified by Age: Late
ASD.AaD_late.path = list.files(here("script_manuscript/genetic_correlation_component/rg_result/"), pattern ="up_late_.*\\.log", full.names = TRUE)
ASD.AaD_late.rg = ldsc_rg_batch_read(ASD.AaD_late.path)  %>% 
  mutate(p.exp_adj = p.adjust(p, method = "fdr"), 
         p.exp_label = case_when(p.exp_adj <0.001 ~ "***", p.exp_adj <0.01 ~ "**", p.exp_adj <0.05 ~ "*", TRUE ~ NA_character_)) %>% 
  rename(GWAS_name = p2_base)

ASD.AaD_late.rg.formatted = sup_table_name_cleaner(ASD.AaD_late.rg) %>%
  sup_table_formatter() %>%
  group_by(MRI_scale, MRI_pheno) %>% # macro-idp = 181; micro-idp = 208; subcor_vol = 26
  mutate(
    p.idp_adj = p.adjust(p, method = "fdr"),
    p.idp_label = case_when(
      p.idp_adj < 0.001 ~ "***",
      p.idp_adj < 0.01  ~ "**",
      p.idp_adj < 0.05  ~ "*",
      TRUE ~ NA_character_)) %>%
  ungroup() %>%
  mutate(p1_base = "Autism_Late_Diagnosed") %>%
  relocate(p, p.label, p.idp_adj, p.idp_label, p.exp_adj, p.exp_label, .after = last_col())

master_rg.df = reduce(list(AQ.rg.formatted, SPQ.rg.formatted, 
                           ASD.rg.formatted, ASD_Daner.rg.formatted, 
                           ASD.AaD_early.rg.formatted, ASD.AaD_late.rg.formatted), full_join)

# save results
write.table(master_rg.df, here("script_manuscript/genetic_correlation_component/master_rg_results.txt"), sep = "\t", quote = FALSE, row.names = FALSE)
```
```{r rg visualisation: Data prep}
# read in and filter data
master_rg.df = fread(here("script_manuscript/genetic_correlation_component/master_rg_results.txt"))

# non-global region visualisation
sig.master_rg.df = master_rg.df %>% 
  filter(MRI_scale == "regional") %>% 
  select(p1_base, MRI_scale, MRI_pheno, region, rg, p.idp_adj, p.idp_label, p.exp_adj, p.exp_label) %>%
    mutate(p_label = case_when(!is.na(p.exp_label) ~ "exp_sig", !is.na(p.idp_label) ~ "idp_sig"),
           p1_base = factor(p1_base, levels = c("AQ_total_score", "SPQ_total_score", 
                                                "Autism_Grove", "Autism_Daner", 
                                                "Autism_Early_Diagnosed", "Autism_Late_Diagnosed")))
```
```{r SF4.1 rg visusalisation: Cortical Brain Regions across all conditions}
rg_reg_brain_vis = sig.master_rg.df %>%
  ggplot(mapping=aes(fill = rg, color = p_label)) + 
  geom_brain(atlas = glasser(), hemi = "left", view = c("lateral", "medial"), size = 0.2) +
  scale_fill_gradient2(low = "#2166AC", mid = "white", high = "coral",
                       midpoint = 0, limits = c(-1, 1), na.value = "grey90", name = "rg") +
  scale_colour_manual(values = c(idp_sig = "black", exp_sig = "red"),na.value = "grey80", name = NULL) +
  facet_grid(MRI_pheno ~ p1_base) +
  theme_void() + 
  guides(colour = "none",
         fill = guide_colourbar(barheight = unit(3, "cm"), barwidth  = unit(0.4, "cm"))) +
  theme(strip.text = element_text(size = 5))

ggsave(here("result_manuscript/SF4.1_rg_brain_plot.svg"), rg_reg_brain_vis, width = 11, height = 8) # To move NA layer level last
ggsave(here("result_manuscript/SF4.1_rg_brain_plot.png"), rg_reg_brain_vis, width = 11, height = 8)
ggsave(here("result_manuscript/SF4.1_rg_brain_plot.pdf"), rg_reg_brain_vis, width = 11, height = 8)
```
```{r F4.3.x significant results}
# Bar plot panel A
sig_global_p1 = master_rg.df %>%
  filter(MRI_scale == "global", p < 0.05) %>%
  pull(p1_base) %>%
  unique()

sig_global_p1.df = master_rg.df %>% 
  filter(p1_base %in% sig_global_p1 & MRI_scale == "global") %>% 
  mutate(MRI_pheno = factor(MRI_pheno, levels = MRI_pheno.list),
         MRI_label = case_when(MRI_pheno %in% DTI_pheno.list ~ "dMRI",
                               MRI_pheno %in% STR_pheno.list ~ "sMRI"),
         p1_name = gsub("_", " ", p1_base))

rg_global_bar_vis = sig_global_p1.df %>%
  ggplot(aes(x = MRI_pheno, y = rg, color = MRI_label)) +
  geom_hline(yintercept = 0, colour = "grey80") +
  geom_point() +
  geom_errorbar(aes(ymin = rg - CI, ymax = rg + CI), width = 0.2) +
  geom_text(aes(label = p.label), vjust = -0.03, size = 4) +
  scale_color_manual(values = c("sMRI" = "coral", "dMRI" = "#2166AC")) +
  coord_flip() +
  ylim(-0.5, 0.5) +
  facet_wrap(~p1_name, scales = "free_y") +
  labs(x = NULL, y = expression(r[g])) +
  theme_classic() +
  guides(color = "none") + 
  theme(strip.text = element_text(size = 10, face = "bold"))

ggsave(here("result_manuscript/F4.3.x_rg_global_plot.png"), rg_global_bar_vis, width = 6, height = 5)
ggsave(here("result_manuscript/F4.3.x_rg_global_plot.pdf"), rg_global_bar_vis, width = 6, height = 5)

# Significant only Panel B
sig_regional_p1 = sig.master_rg.df %>%
  filter(MRI_scale == "regional", p.idp_adj < 0.05) %>%
  distinct(p1_base, MRI_pheno)

sig_regional_p1.df = sig.master_rg.df %>%
  semi_join(sig_regional_p1, by = c("p1_base", "MRI_pheno")) %>%
  filter(MRI_scale == "regional") %>%
  mutate(p1_name = gsub("_", " ", p1_base))

rg_sig.reg_brain_vis = sig_regional_p1.df %>%
  ggplot(mapping=aes(fill = rg, color = p_label)) + 
  geom_brain(atlas = glasser(), hemi = "left", view = c("lateral", "medial"), size = 0.3) +
  scale_fill_gradient2(low = "#2166AC", mid = "white", high = "coral",
                       midpoint = 0, limits = c(-0.5, 0.5), na.value = "grey90", name = "rg") +
  scale_colour_manual(values = c(idp_sig = "black", exp_sig = "red"),na.value = "grey80", name = NULL) +
  facet_wrap(~ p1_name + MRI_pheno, ncol = length(unique(sig_regional_p1.df$p1_base)), dir = "v") +
  theme_void() + 
  guides(colour = "none", fill = guide_colourbar(barheight = unit(3, "cm"), barwidth  = unit(0.4, "cm"))) +
  theme(strip.text = element_text(size = 8))

ggsave(here("result_manuscript/F4.3.x_rg_sig.regional_plot.svg"), rg_sig.reg_brain_vis, width = 14, height = 8) # To move NA layer level last
ggsave(here("result_manuscript/F4.3.x_rg_sig.regional_plot.png"), rg_sig.reg_brain_vis, width = 14, height = 8)
ggsave(here("result_manuscript/F4.3.x_rg_sig.regional_plot.pdf"), rg_sig.reg_brain_vis, width = 14, height = 8)
```
## Statistical Model for Prediction
```{r hubness modelling and graph}
# load network and calculate rg magnitude
rg_network = readRDS(here("script_manuscript/network_clustering_component/rg_network.Rds"))
E(rg_network)$abs_rg = abs(E(rg_network)$rg)

# investigate network properties: Strength, eigenvector centrality, betweenness centrality
net_prop.df = tibble(GWAS_name = V(rg_network)$name, 
                     strength = strength(rg_network, weights = E(rg_network)$abs_rg),
                     eigencent = eigen_centrality(rg_network, weights = E(rg_network)$abs_rg)$vector,
                     betweenness = betweenness(rg_network, weights = 1 / E(rg_network)$abs_rg, normalized = TRUE))

asd_rg.df = master_rg.df %>% 
  filter(p1_base %in% c("Autism_Early_Diagnosed", "Autism_Late_Diagnosed")) %>% # select latent Autism GWAS
  mutate(GWAS_name = paste(MRI_scale, MRI_pheno, region, sep = "_"),
         GWAS_name = gsub("_NA", "", GWAS_name))

asd_net.df = as.data.table(left_join(asd_rg.df, net_prop.df, by = "GWAS_name")) %>% 
  select(p1_base, GWAS_name, MRI_scale, MRI_pheno, region, rg, p, strength, eigencent, betweenness)

ggplot(asd_net.df, aes(strength, p)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm") +
  facet_wrap(~p1_base)

ggplot(asd_net.df, aes(eigencent, p)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm") +
  facet_wrap(~p1_base)

ggplot(asd_net.df, aes(-log(betweenness), p)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm") +
  facet_wrap(~p1_base)
```